### 2017-02-16 csf-2017/paper.tex
Shorten, rephrase.

\documentclass[compsoc,conference,letterpaper,fleqn]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{natbib}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

\usepackage{rotating}
\usepackage{afterpage}

\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\usepackage{enumitem}
\newlist{enumproof}{enumerate}{10}
\setlist[enumproof]{label*=\arabic*.}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{multirow,bigdelim}

\ifCLASSOPTIONcompsoc
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
\usepackage[caption=false,font=footnotesize]{subfig}
\fi

% Include the following to reduce room around display equations
%\setlength{\abovedisplayskip}{3pt}
%\setlength{\belowdisplayskip}{3pt}
%\setlength{\abovedisplayshortskip}{3pt}
%\setlength{\belowdisplayshortskip}{3pt}

% %make latex preview-mode work with natbib...
% \usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\newcommand{\forcenewline}{$\phantom{v}$\\}
\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}

\newcommand{\update}[2]{[#1 \mapsto #2]}
\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}

% Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
\newcommand{\fun}{\rightarrow}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
\renewcommand{\nsim}[1][n]{\stackrel{\tiny{#1}}{\simeq}}

\newcommand\subsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\subset$}\cr
      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
\newcommand\supsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\supset$}\cr
      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
\newcommand{\nsubsim}[1][n]{\stackrel{\tiny{#1}}{\subsetsim}}
\newcommand{\nsupsim}[1][n]{\stackrel{\tiny{#1}}{\supsetsim}}

\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}

\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}
\newcommand{\blater}{\mathop{\blacktriangleright}}
\newcommand{\id}{\var{id}}
\newcommand{\undefined}{\mathit{undefined}}

\newcommand{\powerset}[1]{\mathcal{P}(#1)}

\newcommand{\false}{\mathit{false}}
\newcommand{\true}{\mathit{true}}

% cofes
\newcommand{\cofe}{c.o.f.e.}
\newcommand{\cofes}{\cofe{}'s}
\newcommand{\CatC}{\mathbb{C}}
\newcommand{\CatP}{\mathbb{P}}

% Comments
\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}

% Uncomment me to disable comments
\renewcommand\lau[1]{}
\renewcommand\dominique[1]{}
\renewcommand\lars[1]{}

% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{ms}}
\newcommand{\ms}{\hs}
\newcommand{\hv}{\var{r}}
\newcommand{\rv}{\var{r}}
\newcommand{\lv}{\var{r}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{b}}
\newcommand{\addrend}{\var{e}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{ms}}
\newcommand{\heap}{\var{mem}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
\newcommand{\adv}{\var{adv}}
\newcommand{\msframe}{ms_\var{frame}}
\newcommand{\link}{\var{link}}
\newcommand{\stk}{\var{stk}}
\newcommand{\flag}{\var{flag}}
\newcommand{\nwl}{\var{nwl}}
\newcommand{\pwl}{\var{pwl}}
\newcommand{\sta}{\var{sta}}
\newcommand{\cnst}{\var{cnst}}
\newcommand{\olf}{\var{offsetLinkFlag}}
\newcommand{\prp}{\var{prp}}
\newcommand{\env}{\var{env}}
\newcommand{\cls}{\var{cls}}
\newcommand{\unused}{\var{unused}}
\newcommand{\act}{\var{act}}

% Memory projections
\newcommand{\plainproj}[1]{\mathrm{#1}}
\newcommand{\memheap}[1][\Phi]{#1.\plainproj{mem}}
\newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}

\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{mem}.#2}{#3}}
\newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}

% Configuration end states
\newcommand{\failed}{\mathit{failed}}
\newcommand{\halted}{\mathit{halted}}

% Functions
\newcommand{\plainfun}[2]{
  \ifthenelse{\equal{#2}{}}
  {\mathit{#1}}
  {\mathit{#1}(#2)}
}
\newcommand{\decode}{\plainfun{decode}{}}
\newcommand{\encode}{\plainfun{encode}{}}
\newcommand{\encodePerm}{\mathit{encodePerm}}
\newcommand{\encodePermPair}{\plainfun{encodePermPair}{}}
\newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
\newcommand{\decodePermPair}{\plainfun{decodePermPair}}
\newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}

\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc}{#1}}

\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
\newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}

\newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
\newcommand{\erase}[2]{\floor*{#1}_{\{#2\}}}
\newcommand{\activeReg}[1]{\plainfun{active}{#1}}

% World operations
\newcommand{\future}{\mathbin{\sqsupseteq}}
\newcommand{\pub}{\var{pub}}
\newcommand{\priv}{\var{priv}}
\newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
\newcommand{\futurestr}{\mathbin{\sqsupseteq}^{\var{priv}}}
\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
\newcommand{\memSat}[3][n]{\heapSat[#2]{#1}{#3}}
\newcommand{\memSatPar}[4][n]{\heapSat[#2]{#1 , #4}{#3}}

\newcommand{\monwknefun}{\xrightarrow[\text{\tiny{$\futurewk$}}]{\textit{\tiny{mon, ne}}}}
\newcommand{\monstrnefun}{\xrightarrow[\text{\tiny{$\futurestr$}}]{\textit{\tiny{mon, ne}}}}

% Assembly labels
\newcommand{\codelabel}[1]{\mathit{#1}}
\newcommand{\init}{\codelabel{init}}
\newcommand{\malloc}{\codelabel{malloc}}
\newcommand{\counter}{\codelabel{counter}}
\newcommand{\iocap}{\codelabel{iocap}}

% Type(s)
\newcommand{\type}[1]{\mathrm{#1}}
\newcommand{\asmType}{\plaindom{AsmType}}

% Domains
\newcommand{\plaindom}[1]{\mathrm{#1}}
\newcommand{\Caps}{\plaindom{Cap}}
\newcommand{\Words}{\plaindom{Word}}
\newcommand{\Addrs}{\plaindom{Addr}}
\newcommand{\ExecConfs}{\plaindom{ExecConf}}
\newcommand{\RegName}{\plaindom{RegName}}
\newcommand{\Regs}{\plaindom{Reg}}
\newcommand{\Heaps}{\plaindom{Mem}}
\newcommand{\Mems}{\Heaps}
%\newcommand{\HeapSegments}{\plaindom{MemSegment}}
\newcommand{\HeapSegments}{\plaindom{MemSeg}}
\newcommand{\MemSegments}{\HeapSegments}
\newcommand{\Confs}{\plaindom{Conf}}
\newcommand{\Instrs}{\plaindom{Instructions}}
\newcommand{\nats}{\mathbb{N}}
\newcommand{\ints}{\mathbb{Z}}
\newcommand{\Perms}{\plaindom{Perm}}
\newcommand{\Globals}{\plaindom{Global}}

\newcommand{\Rel}{\plaindom{Rel}}
\newcommand{\Rels}{\plaindom{Rels}}
\newcommand{\States}{\plaindom{State}}
\newcommand{\RegionNames}{\plaindom{RegionName}}
\newcommand{\Regions}{\plaindom{Region}}
\newcommand{\Reg}{\plaindom{Reg}}
\newcommand{\Worlds}{\plaindom{World}}
\newcommand{\Wor}{\plaindom{Wor}}
\newcommand{\Worwk}{\Wor_{\futurewk}}
\newcommand{\Worstr}{\Wor_{\futurestr}}
\newcommand{\xiwk}{\xi_{\var{wk}}}
\newcommand{\xistr}{\xi_{\var{str}}}
\newcommand{\StorePred}{\plaindom{MemSegPred}}
\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
\newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}

\newcommand{\Views}{\plaindom{View}}

% LR
\newcommand{\intr}[2]{\mathcal{#1}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\regintr}[1]{\intr{R}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdrr}{\regintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}
\newcommand{\npairP}[2][n]{(#1,#2)}

% Reference register/memory
\newcommand{\refreg}[1]{#1}
\newcommand{\refheap}[1]{#1}

% Instructions
% No arguments
\newcommand{\zinstr}[1]{\mathtt{#1}}
\newcommand{\fail}{\zinstr{fail}}
\newcommand{\halt}{\zinstr{halt}}
% One argument
\newcommand{\oneinstr}[2]{
  \ifthenelse{\equal{#2}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2}
}
\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
% Two arguments
\newcommand{\twoinstr}[3]{
  \ifthenelse{\equal{#2#3}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2 \; #3}
}
\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
\newcommand{\getl}[2]{\twoinstr{getl}{#1}{#2}}
\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
% Three arguments
\newcommand{\threeinstr}[4]{
  \ifthenelse{\equal{#2#3#4}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2 \; #3 \; #4}
}
\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}
\newcommand{\minus}[3]{\threeinstr{minus}{#1}{#2}{#3}}

% Permissions
\newcommand{\plainperm}[1]{\mathrm{#1}}
\newcommand{\noperm}{\plainperm{o}}
\newcommand{\readonly}{\plainperm{ro}}
\newcommand{\readwrite}{\plainperm{rw}}
\newcommand{\exec}{\plainperm{rx}}
\newcommand{\entry}{\plainperm{e}}
\newcommand{\rwx}{\plainperm{rwx}}
% PWL permissions
\newcommand{\readwritel}{\plainperm{rwl}}
\newcommand{\rwl}{\readwritel}
\newcommand{\rwlx}{\plainperm{rwlx}}

% Global/local
\newcommand{\local}{\plainperm{local}}
\newcommand{\glob}{\plainperm{global}}

\newcommand{\localityReg}{\var{localityReg}}
\newcommand{\localReg}{\var{localReg}}
\newcommand{\globalReg}{\var{globalReg}}

% Views
\newcommand{\plainview}[1]{\mathrm{#1}}
\newcommand{\perma}{\plainview{perm}}
\newcommand{\temp}{\plainview{temp}}
\newcommand{\revoked}{\plainview{revoked}}

% OP sem
\newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
\newcommand{\step}[1][]{\rightarrow_{#1}}

% Conv defs
\newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
\newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
\newcommand{\nonlocal}[1]{\ensuremath{#1} \text{ is non-local}}

% Macros
\newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}

\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
\newcommand\bnfdef{\mathrel{::=}}

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

% domi: IEEEtran.cls sets font size to 10bp, apparently.
% not sure what the difference is with 10pt.
\DeclareMathSizes{10bp}{8}{7}{6}

%\def\IEEEbibitemsep{0pt}

\makeatletter
\setlength\mpr@andskip{1em}
\def\mpr@lineskip{5em}
\def\MathparLineskip{\mpr@lesslineskip}
\makeatother

\newlength{\oldmathindent}
\newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}

\begin{document}
\setlength{\mathindent}{.2cm}

\title{Reasoning about a Capability Machine with Local Capabilities\\
 Provably Safe Stack and Return Pointer Management (without OS Support)}

% \author{%
% \IEEEauthorblockN{Lau~Skorstengaard}
% \IEEEauthorblockA{Aarhus University, Denmark\\
% Email: lau@cs.au.dk} \and
% \IEEEauthorblockN{Dominique~Devriese}
%   \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
% Email: dominique.devriese@cs.kuleuven.be} \and
% \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
% Email: birkedal@cs.au.dk}}
\author{}
\maketitle

\begin{abstract}  
  Capability machines provide security guarantees at machine level which makes
  them an interesting target for compilation schemes that provably enforce
  properties like control-flow correctness, encapsulation of local state. We
  provide a formalization of a representative capability machine with local
  capabilities and propose a novel calling convention for enforcing control-flow
  correctness and encapsulation of local state on it. To prove these properties,
  we provide a logical relation that semantically captures the guarantees
  provided by the hardware (a form of capability safety). These results are not
  tied to our calling convention and can be used to reason about arbitrary
  programs.
\end{abstract}

% \section*{Terminology ?}

% \emph{capability safe}: 
% Our unary logical relation is a semantic \emph{definition}
% of which values for the
% pc / register-file / word are capability safe.

% \emph{adversary}: It doesn't sound ``sexy'' to constantly say ``an untrusted
% piece of code'' should we adopt the convention that adversary refers
% to an unstrusted piece of code?

% \emph{memory}  we don't make a distinction between memory and memory
% segments in the text  (to do so properly, we would need to explain why
% assumptions
% are reasonable, given that a real machine has finite memory, but we
% need infinite domain for malloc spec, etc.)

\section{Introduction}
\label{sec:introduction}

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) or rely on techniques like
machine code rewriting
(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors, intended to
remediate these limitations by presenting a better security model at
the hardware level. They are based on old ideas
(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities, a kind
of pointers carrying authority. Memory capabilities carry
the authority to read and/or write to a range of memory
locations. Moreover, all capability machines
offer some form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

In contrast to the situation for commodity processors, the security
primitives on capability machines lend themselves well to the
enforcement of properties like local-state encapsulation and
control-flow correctness. Potentially, they will enable new compilation
schemes that enforce such properties in a way that is efficient but
also 100\% watertight (ideally evidenced by a mathematical proof,
guaranteeing that we do not end up in a new attack-defense arms
race). However, there is still a lot that needs to happen to attain
this goal. For example, it is quite non-trivial to devise a
compilation scheme adapted to the details of a specific language's
notion of encapsulation (e.g., encapsulation of private member
variables in OO languages often behaves quite differently than
encapsulation of private state in ML-like languages). And even if such
a scheme were defined, a formal proof depends on a formalization of
the encapsulation provided by the capability machine at hand.

A similar problem is the enforcement of control-flow correctness on capability
machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
It drops the standard contiguous C stack and splits it into a central, trusted
shadow stack, managed by trusted call and return instructions, and disjoint,
per-component stacks. To prevent illegal use of stack references, the approach
relies on \emph{local capabilities}, a type of capabilities offered by CHERI
to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
However, a lot of details
remain to be investigated (e.g., how to combine a trusted shadow stack with
(potentially) cross-domain function pointers in a scalable way?) and there is no
argument that this enforcement of control-flow correctness is watertight.

In this paper, we make progress towards the goal of efficient and provably
watertight enforcement of encapsulation and control-flow correctness on
capability machines. Specifically, we make the following contributions:
\begin{itemize}
\item Formalise a simple but representative capability machine featuring local
  capabilities, and its operational semantics
  (Section~\ref{sec:capab-mach-with})
\item Define a novel calling convention enforcing control-flow correctness and
  encapsulation of local state on the stack
  (Section~\ref{sec:stack-and-return-pointer}). It relies solely on local
  capabilities and does not require OS support (like a trusted stack or
  call/return instructions). It also does not use separate per-component stacks
  but a single, fairly standard, contiguous stack. It supports higher-order
  cross-component calls (e.g., (potentially) cross-component function pointers)
  and can be efficient assuming only one additional piece of processor support:
  an efficient instruction for clearing a memory range.
\item Prove a theorem that is the most general and powerful formulation of the
  guarantees offered by a capability machine (a form of capability
  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
  guarantees offered for local capabilities. It is very general and not tied to
  our calling convention or a specific way of using the system's capabilities.
\item Introduce two novel technical ideas in the unary, step-indexed Kripke
  logical relation used to formulate the above theorem. First, we model the
  naturally continuation-passing-style of assembly code (with return pointers as
  continuations) using a \emph{single} orthogonal closure, rather than the
  existing biorthogonal closure. Secondly, we express the special nature of
  local capabilities using a variant of Dreyer et. al.'s public and private
  future worlds~\citep{Dreyer:jfp12}, used to express well-bracketedness
  guarantees in a lambda calculus. The logical relation and capability safety
  theorem is presented in Section~\ref{sec:logical-relation}.
\item Demonstrate our results by applying them to challenging examples,
  specifically constructed to demonstrate local-state encapsulation and
  control-flow correctness guarantees in the presence of cross-component
  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
  the power of our formulation of capability safety and our calling convention.
\end{itemize}
%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution. Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}

For reasons of space, some details and all proofs have been omitted;
refer to the technical appendix~\cite{technical_appendix} for those.

% % Possible high-level intros:
% % 0) Arms race compilers
% %
% % 1) Security guarantees vs obstacles / arms race
% \begin{comment}
%   Security on processors is today subject to an arms race between
%   processor manufacturers and hackers. On the hacker side, a constant
%   effort is put into exploiting current processor designs and
%   circumventing the security measures set in place to prevent known
%   exploits. On the processor manufacturer side, new security measures
%   are put in place to prevent new exploits. The arms race would come
%   to a grinding halt if the security meassures provided more than
%   obstacles and actually provided low-level security guarantees. One
%   kind of low-level machines that provide low-level security is a
%   capability machine.
% \end{comment}
% % 2) Fully abstract compilation
% High-level languages provide guarantees such as encapsulation of local
% state and well-bracketedness of function calls. These guarantees are
% often relied upon when reasoning about the correctness and security of
% a program. High-level programs are compiled to machine code, so in
% order to rely on the correctness and security guarantees provided by
% the high-level language, it needs to be proven that the translation
% preserves these properties. Translated programs often interact with
% machine code that is not translated from the same high-level language,
% so in order to preserve the guarantees from the high-level languages,
% the low-level machine needs to provide some kind of security
% guarantees. In other words, a target of secure compilation needs to
% provide some security guarantees. Modern processors provide many
% security meassures, but very little in terms of guarantees. A machine
% that provides some means of security guarantees is a capability machine.

% % 3) Memory access-control
% % I don't like this angle and can't find a good way to sell it.

% % 4) History
% % Proposals throughout history.
% % Lack of formal account of properties-guarantees.

% %%% a)
% % Current low-level protection coarse grained memory protection or
% % properties of high-level languages are not really enforced?

% % Capability machines offer fine-grained memory protection
% % Too technical
% A capability machine is a low-level machine that offers fine-grained
% memory protection.
% % Capability machines and unforgeable tokens for memory access. 
% This is done by introducing unforgeable tokens to
% the machine. These tokens are called capabillities and grant a kind of
% authority and a range of authority. 
% % Dynamic checks encapsulation, formal model

%  We will refer to the capabilities
% that grant a combination of read, write, or execute permission as memory
% capabilities. 
%  Capabilities with the necessary permissions must be
% presented every time an instruction that manipulates the memory is
% executed.

% Memory capabilities provide memory protection, but they do not provide
% a way to set up security domains. When you jump to an untrusted piece
% of code that is supposed to jump back to you, then you have to either
% revoke all your capabilities or pass to the piece of code you do not
% trust. This short coming can be overcome by adding something like the
% enter capability from the M-Machine\todo{add reference}. The enter
% capability is an opaque capability which can only be used for a
% jump. When jumped to it grants read and execute permission which
% allows the code that is now executing to read capabilities stored in
% memory. 
% The CHERI processor's ccall achieves a similar property\todo{Add reference}.

% Capabilities are irrevocable, so when we pass a capability has to
% an untrusted program, then we have to assume that the program keeps it
% around indefenitely. This means that we cannot reuse the piece of
% memory that the capability governs essentially creating a memory
% leak. The CHERI processor has a special kind of capabilities called
% \emph{local capabilities}\todo{reference?}\ that introduces a simple
% kind of temporal information control which allows for a simple kind of
% capability revocation. This is achieved by adding a tag to every
% capability that marks it as either local or global. Local capabilities
% can only be written through a capability with a new permission called
% \emph{permit write local}.
% % More about local capabilities.

% With just memory capabilities, enter capabailities, and local
% capabilities, the capability machine is very powerful and can
% enforce properties that high-level languages promise. In most
% high-level languages, we expect local state to be private and
% encapsulated. The memory capabilities make sure that local state
% cannot be accessed without a capability to do so. With the enter
% capabaility, we can regain control of local state after passing
% control to an untrusted piece of code which gives us encapsulation
% when control is passed to an untrusted program.

% Many high-level languages guarantee that function function calls are
% well-bracketed. This can easily be achieved by having a trusted call
% stack. On a capability machine, it is possible to enforce well
% barcketedness without a trusted stack. When ensuring well-bracketed
% the two main challenges calls are 1) preventing storage of return
% pointers and 2) encapsulation of the local stack frame. If we don't
% have the first property, then an adversary can store the return
% capability in a call and jump to it in a later call. If the local
% stack frame is not properly encapsulated, then an adversary can break
% the well-bracketedness by jumping to a different calls return
% capability.

% In order to deal with the two challenges, we make sure that all return
% capabilities are local and that there are no global capabilities with
% permit write local permission. This makes sure that there is no way to
% save a return capability in persistent local state.\lau{this does not really make
% sense as the local stack frame can be seen as local state} It is,
% however, to limiting to not provide any means to store the return
% capabilities as a piece of code may need to store a return capability
% before jumping to an adversary. It is therefore needed to provide a
% local capability with permit write local permission which can be used
% as a stack. The stack should now be the only place one can store the
% return capability, but it should also be the only place one can store
% the stack capability.

% Using local return capabilities and putting a stack abstraction on a
% local capability permit write local capability is a good start, but as
% it turns out, it is not wnough. An adversary could do the following 
% \begin{enumerate}
% \item In a first invocation, fill the stack frame with copies of the
%   current stack capability.
% \item In a later call, the adversary may be able to load the stack
%   capability from the previous call.
% \item If we have used a larger part of the stack than in the first
%   call, then the adversary has access to part of our local stack frame.
% \end{enumerate}

% The above attack shows that it is tricky to ensure
% well-bracketedness. It raises the question whether it is possible to
% do so it is completely watertight. The attack utilized the stack, so
% it is necessary to have a trusted stack or can we do without? In
% general it raises the question, how do we reason about capability
% machines and local capabilities, so we are sure that what ever calling
% scheme we come up with actually provably ensures well-bracketed calls.

% In the following paper, we present the following contributions
% \begin{enumerate}
% \item 
% \end{enumerate}

%%% b)

% Examples of capability machines?

% Enter capabilities

% Local capabilities

% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced. Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.

% High-level programs compiled to assembly can ensure these properties.

%

% \subsection{Contributions}

% \begin{itemize}
% \item Formal model of a simple but representative capability machine with local capabilities
% \item Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
% \begin{itemize}
% \item untrusted adversary
% \item higher-order code: callbacks passed to and received from the adversary
% \item efficient stack management
% \item no OS support
% \end{itemize}
% \item Logical relation for reasoning about code in this capability machine
% \item Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
% \item Technical contributions: reuse existing ideas in a new way:
% \begin{itemize}
% \item replace biorthogonal closure by a single orthogonal closure
% because assembly languages remove the distinction between the
% continuation and the arguments
% \end{itemize}
% \end{itemize}
% LB: I think we should have some discussion of the flexibility /
%   strength of our LR.  The LR defines which computations we think of
%   as well-behaved.   
%  We should give simple / trivial examples of
%   code not in the LR. But we may also want to emphasize that the LR is 
%   not too restricting, e.g., it does not enforce a certain calling
%   scheme (examplied by Example \verb!f1! which
%   does not use the stack and the following examples which do use the
%   stack). This ties in to the discussion of biorthogonality, see the
%   following paragraphs.

% LB: we need to relate this carefully to Hur-Dreyer, who used
%       biorthogonality even though they also worked with an assembly
%       language. (If I recall correctly, they assumed some properties
%       of the low level language, which is why the biortho was the
%       right ting, but we need to check.)

% LS: It is my impression that they were interested in the relation
% between high-level and low-level programs. They are therefore in
% particular interested in low-level programs compiled from high-level
% programs which means that the continuation is always invoked in the 
% same way. Our realisation was that anything we pass as an argument 
% in a "call" can be used as a continuation, so the continuation 
% relation was redundant. Even though we do try to give our programs
% some structure using the macros, our logical relation is strong enough
% to handle unstructured programs as well.

% DD: I think the above comment by LS is correct: the reason why the biortho is
% the right thing for Hur-Dreyer is that their LR only accepts well-behaved
% programs which treat the return pointer as a continuation (i.e. don't store
% it), while ours is more general.

% \begin{itemize}
% \item STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
% \end{itemize}
% \begin{itemize}
% \item Demonstrate all of this on several examples:
% \begin{itemize}
% \item security examples
% \item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
% \item some compartmentalisation result?
% \item whatever else we do..
% \end{itemize}
% \end{itemize}

\section{A capability machine with local capabilities}
\label{sec:capab-mach-with}
In this paper, we work with a formal capability machine with all the
characteristics of real capability machines, as well as local capabilities much
like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
infinite address space and unbounded integers.

\begin{figure}
  \begin{align*}
    \addr   &\in& \Addrs &\isdef \nats\\
    w &\in&\Words &\isdef \ints + \Caps \\
    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
    \gl&\in&\Globals & ::= \glob \mid \local \\
     && \Caps  &::= \left\{
                       \begin{multlined}
                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}
                       \end{multlined} \right\}\\
    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
  \end{align*}
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid  \\
           &   & \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid 
                 \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid \\
           &   & \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getp{\lv}{\lv} \mid 
                 \getl{\lv}{\lv} \mid\\ 
           &   & \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlx$};
    \node[main node] (8) [below left of=7] {$\readwritel$};
    \node[main node] (1) [below right of=7] {$\rwx$};
    \node[main node] (2) [below right of=1] {$\exec$};
    \node[main node] (3) [below right of=2] {$\entry$};
    \node[main node] (4) [below left of=1] {$\readwrite$};
    \node[main node] (5) [below right of=4] {$\readonly$};
    \node[main node] (6) [below right of=5] {$\noperm$};

    \path[every node/.style={font=\sffamily\small}]
    (7) edge (8)
    (7) edge (1)
    (8) edge (4)
    (1) edge (2)
    (2) edge (3)
    (2) edge (5)
    (3) edge (6)
    (1) edge (4)
    (4) edge (5)
    (5) edge (6);
  \end{tikzpicture}

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}

\begin{figure*}
  \centering
  \begin{equation*}
    \Phi  \rightarrow
    \begin{cases}
      \sem{\decode(n)}(\Phi) & \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }}\\
      \failed                                 & \text{\footnotesize{otherwise}}
    \end{cases}
  \end{equation*}\vspace{-.3cm}%
  \begin{align*}
    \stdUpdatePc{\Phi} &=
                         \begin{cases}
                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
                             \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}\\
                             \failed & \text{\footnotesize{otherwise}}
                         \end{cases}
  \end{align*}
  \begin{tabular}{|c|p{4.3cm}|p{9.9cm}|}
    \hline
    $i$&$\sem{i}(\Phi)$&Conditions\\
    \hline 
    $\fail$&$\failed$&\\
    \hline
    $\halt$&$(\halted,\memheap)$&\\
    \hline
    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)})$&\\
    \hline
    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$\footnotesize{ and }$\start \leq \addr \leq \addrend$\\
    \hline
    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$\footnotesize{ and }$(\perm',g') = \decodePermPair{\memreg(r_2)}$
                                                      \footnotesize{ and }$(\perm',g') \sqsubseteq (\perm,g)$\footnotesize{ and } $w =((\perm',g'),\start,\addrend,\addr)$\\
    \hline
    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
    \hline
    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline
    \footnotesize{otherwise }$\var{newPc} = \memreg(r)$\\
    \hline
    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$ \footnotesize{ and } $\start \leq \addr \leq \addrend$\footnotesize{ and }$\var{w} = \memreg(r_2)$
                                                                                       \footnotesize{and if }$\var{w} = ((\_,\local),\_,\_,\_)$\footnotesize{, then  }$\perm \in \{\rwlx,\readwritel \}$\\
    \hline
    \multicolumn{3}{|c|}{$\cdots$}\\
    \hline
    \_&$\failed$&\footnotesize{otherwise}\\
    \hline
  \end{tabular}
  \caption{An excerpt from the operational semantics.}
  \label{fig:op-sem}
\end{figure*}

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
We assume an infinite set of addresses $\Addrs$ and define machine words as
either integers or capabilities of the form
$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
authority to execute permissions $\perm$ on the memory range
$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
$\gl$ indicating whether the capability is global or local. There is no notion
of pointers other than capabilities, so we will use the terms interchangeably.
The available permissions and their ordering are depicted in
\figurename~\ref{fig:perm-hier}: the permissions include null permission
($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
three special permissions: read/write-local ($\readwritel$), read/write-local/execute
($\rwlx$) and enter ($\entry$), which we will explain below.

We assume a finite set of register names $\RegName$. We define register files
$\reg$ and memories $\ms$ as functions mapping register names resp. addresses to
words. The state of the entire machine is represented as a configuration that is
either a running $\Phi \in \ExecConfs$ containing a memory and a register file,
or a failed or halted state, where the latter keeps hold of the final state of
memory.
% We also define memory segments $\ms$:
% parts of memory represented as partial functions mapping addresses to words.
%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}

The machine's instruction set is rather basic. Instructions $i$ include
relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
($\move{}{}$, copies words between registers) instructions. Also familiar are
load and store instructions for reading from and writing to memory ($\load{}{}$
and $\store{}{}$) and arithmetic addition operators ($\plus{}{}{}$ and
$\minus{}{}{}$, operating only on numbers). There are three instructions for
modifying capabilities: $\lea{}{}$ (modifies the current address),
$\restricttwo{}{}$ (modifies the permission and local/global tag) and
$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
instructions take care that the resulting capability always carries less
authority than the original (e.g. restrict will only weaken a permission).
Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
$\geta{}{}$ provide access to a capability's permissions, local/global tag, base,
end and current address, respectively.

\begin{comment}
  \begin{figure}
    \centering
    \begin{mathpar}
      \Phi \rightarrow
      \begin{cases}
        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\
          \;\;\text{\footnotesize{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}}\\
          \;\;\text{\footnotesize{and $\perm \in \{ \exec,\rwx, \rwlx \}$} }
        \end{array}\\
        \failed                & \text{\footnotesize{otherwise}}
      \end{cases}
      \and \updatePcPerm{w} =
      \begin{cases}
        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
        w & \text{\footnotesize{otherwise}}
      \end{cases}\\
      \and \stdUpdatePc{\Phi} =
      \begin{cases}
        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
        \begin{array}[t]{l}
          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\ 
          \;\;\text{\footnotesize{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}
        \end{array}\\
        \failed & \text{\footnotesize{otherwise}}
      \end{cases}
      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
      \inferrule{ \memreg(r_2) = \stdcap \\
        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
        \readwritel, \readonly \}\\
        \start \leq \addr \leq \addrend } {
        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
        \updateReg{r_1}{\memheap(\addr)}} \and
      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
        \permp' = \decodePermPair{\memreg(r_2)}\\
        \permp'\sqsubseteq \permp } {
        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
      \inferrule{ \memreg(r_1) = \stdcap\\
        \start \leq \addr \leq \addrend\\
        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
        \var{w} = \memreg(r_2) \\
        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
        \{\rwlx,\readwritel \} } {
        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
    \end{mathpar}
    \caption{An excerpt from the operational semantics.}
    \label{fig:op-sem}
  \end{figure}
\end{comment}
\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics for a
number of representative instructions. Essentially, a configuration $\Phi$ will
either decode and execute the instruction at $\memreg(\pcreg)$ if it is
executable and the current address is in the valid range or fail otherwise. The
table in the figure shows for a given instruction $i$ the result of executing it
in configuration $\Phi$. The $\fail$ and $\halt$ instructions obviously fail and
halt respectively. The $\move{}{}$ instruction simply modifies the register file
as requested and updates the $\pcreg$ to the next instruction using the
meta-function $\stdUpdatePc{}$.

The $\load{}{}$ instruction loads the contents of the requested memory location
into the appropriate register, but only if the capability has appropriate
authority (i.e.\ read permission and an appropriate range). The $\restricttwo{}{}$
instruction will update a capability's permissions and global/local tag in
the register file, but requires that the new permissions are weaker than the
original and does not turn local capabilities into global ones. The $\geta{}{}$
instruction queries the current address of a capability and stores it in another
register.

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their current address and range
cannot be modified. They can only be used to jump to, but when that happens,
their permission is updated to $\exec$. This special behavior makes them
suitable for representing a kind of closures: an opaque package containing a
piece of code together with local encapsulated state. Such a package can be
built as an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where
the range $[\start,\addr-1]$ contains local state (data or capabilities) and
$[\addr,\addrend]$ contains instructions. The package is opaque to an adversary
holding $c$ but when $c$ is jumped to, the instructions can start executing and
have access to the local data through the updated version of $c$ that is then in
$\pcreg$\footnote{Note that the capability in the $\pcreg$-register can be used
  to load from.}.
% \lau{As we
%   talked about what is described here, is not how we make closures in the TR, so
%   we should consider whether we should refrain from using that word here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as all
capabilities with write-local authority are themselves local and the stack is
cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

We point out that our local capabilities capture only a part of the semantics of
local capabilities in CHERI. Specifically, in addition to the above, CHERI's
default implementation of the CCall exception handler forbids local capabilities
from being passed across module boundaries. Such a restriction fundamentally
breaks our calling convention, since we pass around local return pointers and
stack capabilities. However, CHERI's CCall is not implemented in hardware, but
in software, precisely to allow experimenting with alternative models like ours.

% Contrary to more traditional assembly languages, capability machines do not
% allow literal addresses to appear as part of instructions. This means that we

In order to have a reasonably realistic system, we use a simple model of
linking, so programs can have access to other programs through a linking table.
We also assume malloc to be part of the trusted computing base satisfying a
certain specification. Malloc and linking tables are described further in the
next section, but we refer to the technical appendix~\cite{technical_appendix}
for full details.

\section{Stack and return pointer management using local capabilities}
\label{sec:stack-and-return-pointer}
One of the contributions in this paper is a demonstration that local
capabilities on a capability machine support a calling convention that enforces
control-flow correctness in a way that is provably watertight, potentially
efficient, does not rely on a trusted central stack manager, and supports
higher-order interfaces to an adversary. In this section, we explain this
convention's high-level approach, the security measures to be taken in a number
of situations (motivating each separately with a summary table at the end).
After that, we define a number of reusable macro-instructions that can be used
to conveniently apply the proposed convention in subsequent examples.

The basic idea of our approach is simple: we stick to a single, rather standard,
C stack and register-passed return pointers, much like a standard C calling
convention. However, to prevent various ways of misusing this basic scheme, we
put local capabilities to work and take a number of not-always-obvious safety
measures. In the next paragraphs, we will explain the issues to be considered in
all the relevant situations: when (1) starting our program, (2) returning to the
adversary, (3) invoking the adversary, (4) returning from the adversary, (5)
invoking an adversary callback and (6) having a callback invoked by the
adversary.

\textbf{Program start-up} We assume that the language runtime initializes the
memory as follows: a contiguous array of memory is reserved for the stack, for
which we receive a stack pointer in a special register $r_\stk$. We stress that
the stack is not built-in, but merely an abstraction we put on this piece of the
memory. The stack pointer is local and has $\rwlx$ permission. Crucially, the
stack is the only part of memory for which the runtime (including malloc,
loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
Additionally, our examples typically also assume some memory to store
instructions or static data. Another part of memory (called the heap) is
initially governed by malloc and at program start-up, no other code has
capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
assume that memory allocated through malloc cannot be freed.

\textbf{Returning to the adversary} Perhaps the simplest situation is returning
to the adversary after they invoked our code. In this case, we have received a
return pointer from them, and we just need to jump to it as usual. An obvious
security measure to take care of is properly clearing the non-return-value
registers before we jump (since they may contain data or capabilities that the
adversary should not get access to). Additionally, we may have used the stack
for various purposes (register spilling, storing local state when invoking other
functions etc.), so we also need to clear that data before returning to the
adversary.

However, if we are returning from a function that has itself invoked adversary
code, then clearing the used part of the stack is not enough. The \emph{unused}
part of the stack may also contain data and capabilities, left there by the
adversary, including local capabilities since the stack is write-local. As we
will see later, we rely on the fact that the adversary cannot keep hold of local
capabilities when they pass control to the trusted code and receive control
back. In this case, the adversary could use the unused part of the stack to
store local pointers and load them from there after he gets control back. To
prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
the stack that the adversary has had access to, not just the parts that we have
used ourselves. Since we may be talking about a large part of memory, this
requirement is the most problematic aspect of our calling convention for
performance, but see Section~\ref{sec:discussion} for how this might be
mitigated.

\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ to shrink its authority to the part that we are
not using ourselves.

In one of the registers, we also provide a return pointer, which must be a local
capability. If it were global, the adversary would be able to store away the
return pointer in a global data structure (i.e.\ there exists a global capability for it), and jump to it later, in
circumstances where this should not be possible. For example, they could store
the return pointer, legally jump to it a first time, wait to be invoked again
and then jump to the old return pointer a second time, instead of the new return
pointer received for the second invocation. Similarly, they could store the
return pointer, invoke a function in our code, wait for us to invoke them again
and then jump to the old return pointer rather than the new one, received for the
second invocation.
% \lau{If we need space, then maybe one of the examples is sufficient.}
% \lau{Both of these are good to have their as the two last examples of Section~\ref{sec:examples} use them.}
%
% \lau{In the examples, we refer to the local return pointer as
%   ``protected return pointer''. We should either introduce that here
%   or use a different formulation in the examples. }
% \lau{It seemed did not seem to add anything to adopt the terminlogy - we
%   often highlight that the return pointer is local which in most cases
%   is so important that we do not want to leave it out.}
By making the return pointer local, we prevent such attacks: the adversary can
only store local capabilities through write-local capabilities, which (because
of our assumptions above) means: on the stack. Since the stack pointer itself is
also local, it can also only be stored on the stack. Because we clear the part
of the stack that the adversary has had access to before we pass control back,
there is no way for them to recover either of these local capabilities.

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine that the adversary stores his stack pointer, invokes trusted
code that uses part of the stack to store private data and then invokes the
adversary again with a stack pointer restricted to exclude the part containing
the private data. If the adversary had a way of keeping hold of their old stack
pointer, it could access the private data stored there by the trusted code, and
break local-state encapsulation.

\textbf{Returning from the adversary} So return pointers must be passed as local
capabilities. But what should their permissions be, what memory should they
point to and what should that memory (the activation record) contain? Let us
answer the last question first by considering what should happen when the
adversary jumps to a return pointer. In that case, the program counter should be
restored to the instruction after the jump to the adversary, so the activation
record should store this old program counter. Additionally, the stack pointer
should also be restored to its original value. Since the adversary has a more
restricted authority over the stack than the code making the call, we cannot
hope to reconstruct the return pointer from the stack pointer owned by the
adversary. Instead, the old stack pointer should be stored as part of the
activation record.

\begin{figure}
  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
    \draw[draw=gray!80] (0,0.5) -- (0,4);
    \draw[draw=gray!80] (6,0.5) -- (6,4);
    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous program counter};
    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous stack pointer};
    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
    \draw (-3,2.25) node (rp) {return pointer};
    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,2.25);
  \end{tikzpicture}

  \caption{Structure of an activation record}
  \label{fig:activ-rec-struct}
\end{figure}

Clearly, neither of these capabilities should be accessible by the adversary. In
other words, the return pointer provided to the adversary must be a capability
that they can jump to but not read from, i.e.\ an enter capability. To make this
work, we construct the activation record as depicted in
Figure~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
over the entire activation record (containing the previous return and stack
pointer), and its current address points to a number of restore instructions in
the record, so that upon invocation, these instructions are executed and can
load the old stack pointer and program counter back into the register file. As
the return pointer is an enter pointer, the adversary cannot get hold of the
activation record's contents, but after invocation, its permission is updated to
$\exec$, so the contents become available to the restore instructions.

The final question that remains is: where should we store this activation
record? The attentive reader may already see that there is only one possibility:
since the activation record contains the old stack pointer, which is local, the
activation record can only be constructed in a part of memory where we have
write-local access, i.e.\ on the stack. Note that this means we will be placing
and executing instructions on the stack, i.e.\ it will not just contain code
pointers and data. This means that our calling convention should be combined
with protection against stack smashing attacks (i.e.\ buffer overflows on the
stack overwriting activation records' contents). Luckily, such protection is
easy to provide thanks to the capability machine's fine-grained memory
protection.

\textbf{Invoking an adversary callback} If we have a higher-order interface to
the adversary, we may need to invoke an adversary callback. In this case, not so
much changes with respect to the situation where we invoke static adversary
code. The adversary can provide a callback as a capability for us to jump to,
either an $\entry$-capability if they want to protect themselves from us or just
an $\exec$ capability if they are not worried about that. However, there is one
scenario that we need to prevent: if they construct the callback capability to
point into the stack, it may contain local capabilities that they should not
have access to upon invocation of the callback. As before, this includes return
pointers and stack pointers from previous stack frames that they may be
trying to illegally use inside the callback.

To prevent this, we only accept callbacks from the adversary in the form of
global capabilities, which we dynamically check before invoking them (and we
fail otherwise). This should not be an overly strict requirement: our own
callbacks do not contain local data themselves, so there should be no need for
the adversary to construct callbacks on the stack.\footnote{Note that it does
  prevent a legitimate but non-essential scenario where the adversary wants to
  give us temporary access to a callback not allocated on the stack.}

\textbf{Having a callback invoked by the adversary} The above leaves us with
perhaps the hardest scenario: how to provide a callback to the adversary. The
basic idea is that we allocate blocks of memory using malloc that we fill with
the capabilities and data that the callback needs access to, as well as some
prelude instructions that load the data into registers and jumps to the right
code. Note that this implies that no local capabilities can be stored as part of
a closure. We can then provide the adversary with an enter-capability covering
the allocated block and pointing to the contained prelude instructions. However,
the question that remains in this setup is: where do we get a stack pointer when
the callback is invoked?

Our answer is that the adversary should provide it to us, much like we provide
them with a stack pointer when we invoke their code. However, it is important
that we do not just accept any capability as a stack pointer but check that it
is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability. Without
this check, an adversary could potentially get control over our local stack
frame during a callback by passing us a local $\rwx$ capability to a global data
structure instead of a proper stack pointer
and a global callback for our callback to invoke. If our local state contains no
local capabilities, then, following our calling convention, the callback would
not fail and the adversary could use a stored capability for the global data
structure to access our local state. To prevent this from happening, we need to
make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
tells us that the adversary cannot have a global capabilities to our local stack.

\textbf{Calling convention} With the security measures introduced and motivated,
here is a summary of our proposed calling convention:
\begin{description}[font=\normalfont\itshape]
\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure the callback capability is global.
\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} We define a number of reusable macros
capturing the calling convention and other conveniences. All macros that uses
the stack assumes a stack pointer in register $r_\stk$.
\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
$\var{name}$ from the linking table and stores it in register $r$.
\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
invoked by the adversary and prepares the received stack pointer by checking
that it has permission $\rwlx$.
\texttt{\footnotesize{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
capability in register $r$ in the manner described above. That is, it pushes
local state (the contents of registers $\overline{r_{\var{priv}}}$), return
pointer and stack pointer to the stack, sets up the necessary on stack return
code and creates an $\entry$ return pointer, restricts the stack pointer, clears
the unused part of the stack, clears the necessary registers and jumps to $r$.
Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
clears all the memory the capability in register $r$ has authority over.
\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
register $r$ is a global capability. \texttt{\footnotesize{crtcls
    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) where the
contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

The technical appendix~\cite{technical_appendix} contains detailed descriptions
of all the macros.

\section{Logical relation}
\label{sec:logical-relation}
In order to reason about programs on the capability machine and prove properties
such as well-bracketedness and encapsulation, we define a unary logical
relation.
% Intuition
It is a semantic model of all the words that are definitely safe to give to an
adversary. Whether a capability is safe depends heavily on the contents of the
memory since instructions as well as capabilities may reside there. Hence we
index the logical relation over worlds, modelling the memory (this makes our
logical relation into a Kripke logical relation). A world consists of a finite
set of regions, and each region defines a protocol that governs the possible
evolution of a part of memory. A future world relation defines how worlds may
evolve during execution (e.g. new protocols established over memory received
from malloc, or steps in a protocol already established for a part of memory).
In fact, as we will explain below, there are two future world relations to
accommodate reasoning about local capabilities.
%

\subsection{Worlds}
\begin{figure}
  \centering
  \includegraphics{w11}
  \caption{An illustration of local/global capabilities and temporary/permanent
    regions. The table represents a memory and the colored fields are regions
    governing a memory range. Global capabilities cannot depend on temporary
    regions.}
  \label{fig:cap-world}
\end{figure}
% \dominique{Perhaps we could first try to convey the intuition of a permanent
%   world with a temporary overlay, with global capabilities ignoring the overlay.
%   Perhaps we could add a depiction of the world to Figure 5?}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, to describe a protocol for the region's memory,
similarly to what has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
imposed by permanent regions stay in place indefinitely. Any capability, local
or global, can depend on these protocols. Protocols imposed by temporary regions
can be revoked at certain points during execution, on the condition that we
clear local capabilities when this happens. This means that local capabilities
can safely depend on the protocols imposed by temporary regions, but global
capabilities cannot, since a global capability may outlive a temporary region
that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.

For technical reasons, we do not actually remove a revoked temporary region from
the world, but we turn it into a special revoked region that exists for this purpose.
Such a revoked region contains no state transition system and puts no
requirements on the memory. It simply serves as a mask for a revoked temporary
region. Masking a region like this goes back to earlier work of
\citet{Ahmed2004semantics} and was also used by
\citet{Thamsborg:2011:KLR:2034773.2034831}.

% Recursive worlds intuition
% The worlds end up being defined recursively: The memory can contain
% capabilities,
% which grant authority over parts of memory. A world
% models memory, so it also needs to be able to model capabilities in
% memory and the protocols on them. Whether a capability satisfies the
% protocol of a region may depend on what it grants authority over in
% memory which is described by the world.

%\dominique{``end up being defined recursively'' does not sound very insightful}
A region defines what its part of memory can legally look like. However,
validity of the memory's contents may itself depend on the state of other parts
of memory. To accommodate this, worlds need to be defined recursively: Worlds
describe memory, which can contain capabilities, whose safety is
world-dependent. Recursive worlds are a common requirement in Kripke models and
the following lemma uses the method of \citet{Birkedal:2011:SKM:1926385.1926401}
for constructing them. The formulation of the lemma is technical, so we
recommend that non-expert readers ignore the technicalities and accept that
there exists a set of worlds $\Wor$ and two relations $\futurestr$ and
$\futurewk$ satisfying the (recursive) equations in the theorem (where the
$\blater$ operator can be safely ignored).
% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
% \dominique{it is not explained what cofes are?}
% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  \begin{align*}
      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions)\\
      \Regions & = \{\revoked\} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\temp\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
                   \end{multlined} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\perma\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
                   \end{multlined}
    \end{align*}
  and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\[
  \Worlds = \nats \finparfun \Regions
\]

\subsubsection{Future worlds}
The future world relations model how memory may evolve over time. The
\emph{public future world} $W' \futurewk W$ requires that $\dom(W') \supseteq
\dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$. That is, in a
public future world, new regions may have been allocated, and existing regions
may have evolved according to the public future region relation (defined below).
The \emph{private future world} relation $W' \futurestr W$ is defined similarly,
using a private future region relation.

The \emph{public future} region relation is the simplest. It satisfies
% \dominique{is defined by?} 
% \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations. It can be shown that these will have the properties mentioned below}
the following properties:
\begin{mathpar}
  \inferrule{ (s,s') \in \phi_\pub }
  {  (v,s',\phi_\pub,\phi,H) \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
Both temporary and permanent regions are only allowed to transition according to
the public part of their transition system. Additionally, revoked regions must
either remain revoked or be replaced by a temporary region. This means that the
public future world relations allows us to reuse a region that has been revoked
earlier.

The \emph{private future region} relation satisfies:
\begin{mathpar}
  \inferrule{ (s,s') \in \phi} 
  {  (v,s',\phi_\pub,\phi,H) \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
Here, revocation of temporary regions is allowed. As explained above, this can
be done by replacing them with a special revoked region, which does not describe
any memory, but in fact, any other region can also be used to replace both
revoked and temporary regions. On the other hand, permanent regions cannot be
masked away. They are only allowed to transition according to the private part
of the transition system.

Notice that the public future region relation is a subset of the
private future region relation.

%%%%%%%%

% In order to model changes to memory over time, we use the future world
% relations. The future world relations model how more memory may be
% allocated but also how memory may change. How memory is allowed to
% change over time depends on the region that governs it and the future
% world relation used. The most interesting part about the future worlds
% is how it affects the regions, so we first look at the properties of
% the \emph{public future region} relation and the \emph{private future
%   region} relation. We then use these relations to describe the future
% world relations.

% {\bf Private future regions}
% In the private future region relation, we allow revocation of $\temp$
% regions. The regions cannot be revoked by removing them from the
% world, because we need the future world relations to be
% monotone. This is why we have the $\revoked$ region, so we have
% a region that can serve as a mask. We can, however, let any region
% serve as a mask, so in the private future region relation, we allow a
% $\temp$ region to become any valid region. As the $\revoked$ region
% serves as a mask, we let any valid region take its place in the
% private future region relation.

% If a $\temp$ region is not masked by a region, then it behaved like a
% $\perm$ region. The $\perma$ regions stay the same type and they are
% allowed to take a transition in the private part of the transition
% system. Which can be thought of as the memory changes according to
% some protocol.

% The properties of the \emph{private future region} relation are:
% \begin{mathpar}
%   \inferrule{  (s,s') \in \phi \\
%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
%   {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ r \in \Regions }
%   { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ r \in \Regions }
%   { r \futurestr \revoked }
% \end{mathpar}

% {\bf Public future regions}
% In the public future regions we do not permit revocation of $\temp$
% regions. In this relation, the $\temp$ and $\perma$ regions must stay
% the same type and they are allowed to take a transition in the public
% part of the transition system.

% The $\revoked$ regions serve as masking regions, but are not as
% liberal as in the private future region relation when it comes to
% letting other regions replace them as masks. In the public future
% region relation, we only allow $\revoked$ and $\temp$ regions to take
% on the job as masking region. We have this limitation, because we
% want the private future region relation to be able to reuse any region
% it has revoked.

% The \emph{public future} region relations satisfies the following properties:
% \begin{mathpar}
%   \inferrule{  (s,s') \in \phi_\pub \\
%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
%   {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
%   \and
%   \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
%   { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
%   \and
%   \inferrule{ }
%   { \revoked \futurewk \revoked }
% \end{mathpar}

% Notice that the public future region relation is a subset of the
% private future region relation, so anything you are allowed to do in
% the public future region relation, the private future region relation
% can do the same.

% % The future world relation are really the ones from thm, the above
% % are properties of them.
% {\bf Future world relations}
% There are two future world relations a \emph{public future world}
% relation and a \emph{private future world} relation. In both future
% world relations, we have extension ordering, so a future world $W'$ of
% $W$ has at least the same regions as $W$, but $W'$ may have new regions
% imposing new protocols on parts of the memory. The two future world
% relations differ in the way existing regions are allowed to
% evolve. The public future world relation uses the public future region
% relation and the private future world relation uses the private future
% region relation:
% \begin{mathpar}
%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
%     \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
%   { W' \futurewk W }
%   \and
%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
%     \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
%   { W' \futurestr W }
% \end{mathpar}

% % Relate to STSs for high-level languages. Have the public-private
% % transitions, but they are also used for more than this
% \todo[inline]{Relate to STSs for high-level languages?}

\subsubsection{World satisfaction}
A memory satisfies a world, written $\memSat{\ms}{W}$, if it can be partitioned
into disjoint parts such that each part is accepted by an active (permanent or
temporary) region.
\begin{align*}
  &\memSat{\ms}{W}
    \text{ iff }\\
  &\quad \left\{\begin{aligned}
        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\[-.2cm]
        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
      \end{aligned}\right.
\end{align*}

\subsection{Logical relation}
\afterpage{
\begin{figure}[htbp]
  \centering
  \begin{withmathindent}{0cm}
    \begin{align*}
      & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
      & \observations (W) \defeq \\
      & \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
        \begin{aligned}
          & \forall \ms_f, \heap', i \leq n \ldotp \\
          & (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
          & \;\; \Rightarrow
          \begin{aligned}[t]
            & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{\footnotesize{ and }} \\
            & \;\; \heapSat[\hs']{n-i}{W'}
          \end{aligned}
        \end{aligned}\right\}\\
      & \stdrr \; : \; \Worlds \monwknefun \UPred{\Regs} \\
      & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
          \forall r \in \RegName \setminus \{\pcreg\} \ldotp \npair{\reg(r)}
          \in \stdvr(W) \right\} \\
      & \\[-.2cm]
      & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
      & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \;
        \begin{multlined}
          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
          \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W)
        \end{multlined} \right\} \\
      & \\
      &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
      &\stdvr(W)\defeq  \\
      &\;\;\begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \}
        \union \\
        & \{ \npair{\stdcap[(\noperm,\gl)] } \}
        \union \\
        & \left\{
          \begin{multlined}
            \npair{\stdcap[(\readonly,\gl)] } \mid \\
            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
          \end{multlined} \right\}
        \union \\
        & \left\{
          \begin{aligned}
            &\npair{\stdcap[(\readwrite,\gl)] } \mid \\
             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
             &\quad\npair{(\start,\addrend)} \in
            \writeCond{}(\iota^\nwl,\gl)(W)
          \end{aligned} \right\} \union \\
        & \left\{
          \begin{aligned}
            &\npair{\stdcap[(\readwritel,\gl)] } \mid \\
             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
             &\quad\npair{(\start,\addrend)} \in
            \writeCond{}(\iota^\pwl,\gl)(W)
          \end{aligned} \right\}
        \union \\
        & \left\{
          \begin{aligned}
            &\npair{\stdcap[(\exec,\gl)]} \mid \\
             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
             &\quad\npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
          \end{aligned} \right\}
        \union \\
        & \left\{
          \begin{multlined}
            \npair{\stdcap[(\entry,\gl)]} \mid \\
            \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)
          \end{multlined} \right\}
        \union \\
        & \left\{
          \begin{aligned}
            &\npair{\stdcap[(\rwx,\gl)]} \mid \\
            &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
            &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
            &\quad\npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
          \end{aligned} \right\}
        \union \\
        & \left\{
          \begin{aligned}
            &\npair{\stdcap[(\rwlx,\gl)]} \mid \\
             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
             &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
             &\quad\npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
            \execCond{}(\gl)(W)
          \end{aligned} \right\}
      \end{aligned}
    \end{align*}
  \end{withmathindent}
  \caption{Logical Relation}
  \label{fig:logrel}            
\end{figure}

% \begin{figure}[htbp]
%   \centering
%   \begin{align*}
%   &\iota_{\start,\addrend}^\pwl : \Regions\\
%   &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
%   \\
%   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
%   &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
%     \begin{aligned}
%       & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
%       &\forall \addr \in [\start,\addrend] \ldotp\\
%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
%     \end{aligned}
%         \right\}\\
%   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
%   &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
%   \\
%   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
%   &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
%     \begin{aligned}
%       &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
%       &\forall \addr \in [\start,\addrend] \ldotp \\
%       & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
%      \end{aligned}
%      \right\}
% \end{align*}
% \caption{Standard regions}
% \label{fig:standard-regions}
% \end{figure}

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation now defines semantically when values, terms, and
configurations can safely be used or executed by the adversary, i.e. without the
risk that they will be able to use it to break protocols in the world. The
definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
and~\ref{fig:perm-conds} and we provide some explanations in the following
paragraphs. We omit some definitions and explain them only verbally, but precise
definitions can be found in the technical appendix~\cite{technical_appendix}.

The \emph{observation relation} $\observations$ defines configurations that we
consider safe. Roughly speaking, a configuration is safe if whenever it halts,
there is a private future world that the resulting memory satisfies. Notice that
failing is considered safe behavior. In fact, the machine often resorts to
failing when an unauthorized access is attempted, like loading from a capability
without read permission. This is similar to \cite{Devriese:2016ObjCap}'s logical
relation for an untyped language, but unlike typical logical relations
for typed languages, which require that programs do not fail.

The \emph{register-file relation} $\stdrr$ specifies that register-files are
safe when they contain safe words (i.e.\ words in $\stdvr$) in all registers but
$\pcreg$. The \emph{expression relation} $\stder$ defines that a word is safe to
use as a program counter if it can be plugged into a safe register file (i.e.\ a
register file in $\stdrr$) and paired with a memory satisfying the world to
produce a safe configuration.

The \emph{value relation} $\stdvr$ defines when words are safe, i.e.\ when they
cannot be used to break the protocols defined by the world. Non-capability data
is always safe because it provides no authority. Capabilities give the authority
to manipulate memory and potentially break memory protocols, so they need to
satisfy certain conditions to be safe. In \figurename~\ref{fig:perm-conds}, we
define such a condition for each kind of permission a capability can have (we
will call these permission-based conditions).

For capabilities with read permission, the $\readCond{}$ ensures that it can
only be used to read safe words, by requiring that the region governing the
relevant part of memory, guarantees that it contains only words in the value
relation. Generally speaking, global capabilities should only depend on
permanent regions. Therefore we use the function $\var{localityReg}(\gl,W)$,
which projects out all active regions when $\gl$ (the locality) is local, but
only the permanent regions when $\gl$ is global. The read condition is defined
in terms of a standard region $\iota^\pwl_{\start,\addrend}$ which requires all
the words in the range $[\start,\addrend]$ to be safe.

For a capability with write permission, $\writeCond{}$ must be
satisfied for the capability's range of authority. Such a capability
may be used to write any word, so the region governing the relevant
memory must (at least) allow any safe word to be stored. More
precisely, for write-local capabilities it needs to allow arbitrary
safe values to be stored, while for write capabilities, storing values
that are both non-local and safe is sufficient.  This is done in the
value relation by the regions the write condition is applied with. For
write-local capabilities $\iota^\pwl$ is used, and for write
capabilities we use another standard region
$\iota^\nwl_{\start,\addrend}$ which requires that the words in
$[\start,\addrend]$ are non-local and safe.  Like before, we use
$\var{localityReg}$ to pick an appropriate region for the capability's
locality. The condition for the region to be address-stratified is a
technical requirement, defined in the technical appendix.

The conditions $\entryCond{}$ and $\execCond{}$ are very similar. Both require
that the capability can be safely jumped to. However, executable capabilities
can be updated to point anywhere in their range, so they must be safe as a
program counter (in the $\stder$-relation) no matter where they point to. In
contrast, enter capabilities are opaque and can only be used to jump to the
address they point to. They also change permission when jumped to, so we require
them to be safe as a program counter when the permission is changed to $\exec$.
Because the capabilities are not necessarily invoked immediately, this must be
true in any future world, but it depends on the locality of the capability which
future worlds we consider. If it is global, then it should only depend on
permanent regions, so we require safety as a program counter in \emph{private}
future worlds (recall that temporary regions may be revoked in a private future
world). For local capabilities, it suffices to be safe in \emph{public} future
worlds, where temporary regions are still present.

\subsection{Safety of the capability machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalises the guarantees offered by
the capability machine. Essentially, it says that a capability that only grants
safe authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  If one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
The Fundamental Theorem can be understood as a general expression of the
guarantees offered by the capability machine, an instance of a general property
called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
understand this, consider that the theorem says the capability
$((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter, without any
assumption about what instructions it actually points to (the only assumptions
we have are about the read or write authority that it carries). As such, the
theorem expresses the capability safety of the machine, which guarantees that
\emph{any} instruction is fine and will not be able to go beyond the authority
of the values it has access to.

From another point of view, the Fundamental Theorem also serves as a sanity
check of the logical relation as it guarantees that a large class of
capabilities actually inhabits it.

\section{Examples}
\label{sec:examples}
In this section, we demonstrate how our formalisation of capability safety
allows us to prove local-state encapsulation and control-flow correctness
properties for challenging program examples. The security measures of
Section~\ref{sec:stack-and-return-pointer} are deployed to ensure these
properties. Since we are dealing with assembly language, there are many details
to the formal treatment, and therefore we necessarily have to omit some details
in the lemmas we state here. The interested reader can find all the technical
details in the technical appendix~\cite{technical_appendix}.

\subsection{Encapsulation of local state}
\texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's
encapsulation of local state. They are very similar: both store some local
state, call an untrusted piece of code ($\adv$), and then test whether the local
state is unchanged. They differ in the way they do this. Program
\texttt{\footnotesize{f1}} assumes access to a stack pointer and uses the stack
to store its local state and the stack-based calling convention
\texttt{\footnotesize{scall}} to call the adversary. On the other hand,
\texttt{\footnotesize{f2}} uses malloc to allocate memory for its local state
and uses an activation-record based calling convention (described in the
technical appendix) to run the adversarial code.

\begin{figure}[t]
  \centering

  \begin{minipage}[t]{4.1cm}
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
  \end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{4.1cm}
  \begin{lstlisting}
f2: malloc $r_l$ 1
    store $r_l$ 1
    fetch $r_1$ $\adv$
    call $r_1$($[]$,$[r_l]$)
    assert $r_l$ 1
    halt
  \end{lstlisting}
  \end{minipage}
  \caption{Two example programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} does not rely on a stack.}
  \label{fig:prog-f1-and-f2}
\end{figure}

For both programs, we can prove that if they are linked with an adversary,
$\adv$, that is allowed to allocate memory but has no other capabilities, then
the assertion will never fail during executing (see
Lemmas~\ref{lem:correctness-f1} and~\ref{lem:correctness-f2} below).
% \lars{shouldn't we say already
% here that the adversary is a global capability and that is why,
% intuitively, the local state is preserved ?} \lau{Due to the ``order
% of control'', \texttt{f1} runs first. The lemma is quite restrictive
% on the adversary in the sense that it only has access to malloc (and
% whatever we give et access to). Specifically, it does not have access
% to the stack pointer and because of that we don't need to check it
% first. We could write that the lemmas are limited in their assumptions
% to make them simpler and a more complicated example is coming up.}
The two examples also illustrate the versatility of the logical relation.
Because it is not specific to any calling convention, we can use it to reason
about both programs, even though they use different calling conventions.

% \begin{lemma}[Correctness lemma for \texttt{f1}]
% \todo[inline]{The lemma copied directly from the TR. To be deleted later!}
%   \label{lem:correctness-f1}
%   Let
%   \begin{align*}
%     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
%     c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
%     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
%     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
%     c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
%     \reg & \in \Regs \\
%     m & \defeq \hs_{f1} \uplus 
%         \hs_\flag \uplus                
%         \ms_{\var{link}} \uplus 
%         \hs_\adv \uplus 
%         \ms_{\malloc} \uplus 
%         \ms_{\var{stk}} \uplus
%         \ms_{\var{frame}} 
%   \end{align*}
%   and
%   \begin{itemize}
%   \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
%   \end{itemize}
%   where 
%   \begin{align*}
%     &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
%     &\dom(\hs_\flag) = [\flag,\flag] \\
%     &\dom(\ms_\link) = [\link,\link+1]\\
%     &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
%     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
%     &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
%   \end{align*}
%   and
%   \begin{itemize}
%   \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
%   \item $\ms_\flag = [\flag \mapsto 0]$
%   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
%   \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
%   \end{itemize}
%   if 
%   \[
%     (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
%   \]
%   then
%   \[
%     m'(\flag) = 0
%   \]  
% \end{lemma}

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory)
that they set to $1$ if the assertion fails and then halt. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is safe (in the $\observations$ relation) for a world with a
permanent region that requires the assertion flag to be 0. By an anti-reduction
lemma, it suffices to show that the configuration is safe after some reduction
steps. We then use a general lemma for reasoning about
\texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
configuration that \texttt{\footnotesize{scall}} will jump to is safe and (2)
that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
up is safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the adversary capability is an enter capability, which the
Fundamental Theorem says nothing about. Luckily the enter capability becomes
$\exec$ after the jump and then the Fundamental Theorem applies.

We have a similar lemma for \texttt{\footnotesize{f2}}:
\begin{lemma}
  \label{lem:correctness-f2}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer,
  if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{\footnotesize{f3}} and
\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-f3-and-g1}.

In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order
for the assertion in the middle to succeed, they need to be well-bracketed. If
the adversary were able to store the return pointer from the first call and
invoke it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on
top of its stack and the assertion would fail. However, the security measures in
Section~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
the return pointer is local, so it can only be stored on the stack, but the part
of the stack that is accessible to the adversary is cleared before the second
invocation. In fact, the following lemma shows that there are also no other
attacks that can break well-bracketedness of this example, i.e.\ the assertion
never fails. It is similar to the two previous lemmas:

\begin{comment}
\newsavebox{\testit}
\begin{lrbox}{\testit}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
\end{lrbox}

\newsavebox{\progfthree}
\begin{lrbox}{\progfthree}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
\end{lrbox}

\begin{figure}[t]
  \centering
  \subfloat{\usebox{\testit}}
  \subfloat{\usebox{\progfthree}}
  \caption{test}
\end{figure}
\end{comment}
\begin{figure}[t]
  \centering
  \begin{minipage}[t]{5.05cm}
  \begin{lstlisting}
g1:  malloc $r_2$ 1
     store $r_2$ 0
     move $\pcreg$ $r_3$
     lea $r_3$ $\var{offset}$
     crtcls $[(x, r_2)]$ $r_3$
     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
     jmp $r_0$
f4:  reqglob $r_1$
     prepstk $r_\stk$
     store $x$ 0
     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
     store $x$ 1
     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
     load $r_1$ $x$
     assert $r_1$ 1
     mclear $r_\stk$
     rclear $\RegName \setminus \{r_0,\pcreg\}$
     jmp $r_0$
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{3.4cm}
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \end{minipage}
  \caption{ Two programs that rely on well-bracketedness of
    \texttt{scall}s to function correctly. $\var{offset}$ is the
    offset to \texttt{f4}.}
  \label{fig:prog-f3-and-g1}
\end{figure}

\begin{lemma}
  \label{lem:correctness-f3}
  Making similar assumptions about capabilities and linking as in
  Lemma~\ref{lem:correctness-f1}
  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
  (\halted,m')$, then $m'(\flag) = 0$.
\end{lemma}

The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
is a faithful translation of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
\texttt{\footnotesize{g1}} is a closure generator that generates closures with
one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
the program (note that we can omit some calling convention security measures
because the stack is not used or passed to the adversary in the closure
generator). \texttt{\footnotesize{f4}} expects one argument, a callback. It sets
$x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
the callback a second time. When it returns again, it asserts that $x$ is $1$
and returns. This example is more complicated than the previous ones because it
involves a closure invoked by the adversary and an adversary callback invoked by
us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
need to check (1) that the stack pointer that the closure receives from the
adversary (it must have the correct permission and fit with our calling
convention) and (2) that the adversary callback is global.

To illustrate how subtle this program is, consider how an adversary could try to
make the assertion fail. In the second callback an adversary can get to the
first callback by invoking the closure one more time. If there were any way for
the adversary to transfer the return pointer from the point where it reinvokes
the closure to where the closure reinvokes the callback, then the assertion
could be made to fail. Similarly, if there were any way for the adversary to
store a stack pointer or trick the trusted code into preserving it across an
invocation, the assertion can likely be made to fail too. However, our calling
convention prevents any of this from happening, as we prove in the following
lemma.

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if
  \[
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
  \]
  then $m'(\flag) = 0$.
\end{lemma}

As explained in Section~\ref{sec:stack-and-return-pointer}, the
macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
is global, essentially to make sure it is not allocated on the stack where it
might contain old stack pointers or return pointers. Otherwise, the
encapsulation of our local stack frame could be broken. In the proof of
Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
callback in a world that is only a private future world of the one where we
received the callback, precisely because we have invalidated the adversary's
local state (particularly their old stack and return capabilities). The callback
is still valid in this private future world, but only because we know that it is
global.

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{\footnotesize{g1}}. This
means we need to check that all arguments are safe to use and that we clean up
before returning in the end. The inversion of control makes it challenging to
reason about the adversary's local state during the execution of
\texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}}
and the callbacks, the adversary cannot use their local state from before the
call of \texttt{\footnotesize{f4}}. This is modeled by revoking the temporary
regions of the world we get at the start of \texttt{\footnotesize{f4}}. However,
when \texttt{\footnotesize{f4}} returns, the adversary's old local state becomes
available again, since they are allowed to allocate their activation record on
the stack (just like we do) so that they can store and recover their old stack
pointer after \texttt{\footnotesize{f4}} returns. This means that we need to
restore validity of the adversary's old local state, which we achieve
essentially by clearing the stack.

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
the proof of Lemma~\ref{lem:correctness-g1} by constructing suitable worlds for
all situations where we pass control to the adversary. The potentially local
return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe
in $W_1$, so it can only be used in public future worlds. Let us see how $W_6$
is constructed as a private future world of $W_5$ but also a public future world
of $W_1$. The worlds $W_2$ and $W_4$ are constructed by revoking all temporary
regions and adding a $\iota^\pwl$ region for the stack we pass away and static
temporary regions (which accept exactly one memory) for our local stack as well
as the adversary's old local stack. The given worlds $W_3$ and $W_5$ are public
future worlds in which the adversary chooses to return to us. None of these
worlds can have replaced the temporary regions of $W_1$ with permanent regions.
Therefore, by revoking the temporary regions of $W_5$ and reinstating the
temporary regions of $W_1$, we can construct $W_6$ as a private future world of
$W_5$ but also a public future world of $W_1$. Because of the latter, it is safe
to use the local return pointer to return to the adversary, since we have
restored validity of any local state they might have stashed away.

\begin{figure}[t]
  \centering
\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
  \node[main node] (1) {$W_1$};
  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f3} called}}};
  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};

  \node[main node] (2) [above right of=1]{$W_2$};
  \node[main node] (label2) [above of=2,yshift=-1cm] {\textit{\footnotesize{first callback}}};

  \node[main node,left] (const) [above of=given,yshift=-0.4cm] {\parbox{1.5cm}{\footnotesize{Constructed:}}};

  \node[main node] (3) [below right of=2]{$W_3$};
  \node[main node] (label3) [below of=3,yshift=0.75cm] {\footnotesize{\textit{callback returns}}};

  \node[main node] (4) [above right of=3]{$W_4$};
  \node[main node] (label4) [above of=4,yshift=-0.75cm] {\footnotesize{\textit{second callback}}};

  \node[main node] (5) [below right of=4]{$W_5$};
  \node[main node] (label5) [below of=5,yshift=1cm] {\footnotesize{\textit{callback returns}}};

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
\caption{Illustration of the worlds in the proof of
  Lemma~\ref{lem:correctness-g1}. In the proof, the above row of worlds are
  constructed by us, while the below worlds are given. $W_6$ is constructed such
  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{comment}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?

For example, an untrusted advertisement scenario with initialization code
that registers a redraw callback. The redraw callback gets temporary
read-write access to a framebuffer.

\item Some compartmentalisation result?
\end{itemize}
\end{comment}

\section{Discussion}
\label{sec:discussion}
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling
convention enforces control-flow correctness, we do not prove a general theorem
that shows this, because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by
\cite{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
In fact, our examples show that our logical relation imply a stronger form of
control-flow correctness than such formulations, although this is not made very
explicit. As future work, we consider looking at a more explicit and useful way
to formalise control flow correctness. The idea would be to define a variant
of our capability machine with call and return instructions and
well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our calling convention is
fully abstract~\citep{abadi_protection_1998}.

\emph{Performance and the requirement for stack clearing} The additional
security measures of the calling convention described in
Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
calling convention without security guarantees. However, most security measures
require only a few atomic checks or register clearings on boundary crossings
between trusted code and adversary, which should produce an acceptable
performance overhead. The only exception are the requirements for stack clearing
that we have in two situations: when returning to the adversary and when
invoking an adversary callback. As we have explained, we need to clear all of
the stack that we are not using ourselves, not just the part that we have
actually used. In other words, on every boundary cross between trusted code and
adversary code, a potentially large region of memory must be cleared. We believe
this is actually a common requirement for typical usage scenarios of local
capabilities and capability machines like CHERI should consider to provide
special support for this requirement, in the form of a highly-optimized
instruction for erasing a large block of memory.

\emph{Modularity} It is important that our calling convention is modular, i.e.
we do not assume that our code is specially privileged w.r.t. the adversary, and
they can apply the same measures to protect themselves from us as we do to
protect ourselves from them. More concretely, the requirements we have on
callbacks and return pointers received from the adversary are also satisfied by
callbacks and return pointers that we pass to them. For example, our return
pointers are local capabilities because they must point to memory where we can
store the old stack pointer, but the adversary's return pointers are also
allowed to be local too. Adversary callbacks on the other hand are required to
be global but the callbacks we construct ourselves are allocated on the heap and
also global. \\\\
\noindent\textbf{Logical relation}\\
\emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
Figure~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
existing pattern called biorthogonality. Biorthogonality is a pattern for
defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998}
in terms of an observation relation of safe configurations (like we do). The
idea is to define safe evaluation contexts as the set of contexts that produce
safe observations when plugging safe values and define safe terms as the set of
terms that can be plugged into safe evaluation contexts to produce safe
observations. This is an alternative to more direct definitions where safe terms
are defined as terms that evaluate to safe values. An advantage of
biorthogonality is that it scales better to languages with control effects like
call/cc. Our definitions can be seen as a variant of biorthogonality, where we
take only a single orthogonal closure: we do not define safe evaluation contexts
but immediately define safe terms as those that produce safe observations when
plugged with safe values. This is natural because we model arbitrary assembly
code that does not necessarily respect a particular calling convention. As a
result, return pointers are in principle values like all others and there is no
reason to treat them specially in the logical relation.

Interestingly, \citet{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
Kripke logical relation for an assembly language (for reasoning about correct
compilation from ML to assembly), but because they only model non-adversarial
code that treats return pointers according to a particular calling convention,
they can use standard biorthogonality rather than a single orthogonal closure
like us.

\emph{Public/private future worlds} A novel aspect of our logical relation is
how we model the special nature of local capabilities using public/private
future worlds. The main insight is that the temporary, revokable nature of these
capabilities can be seen as a generalization of the syntactically-enforced
unstorable status of evaluation contexts in lambda calculi without control
effects. To reason about code that relies on this (particularly, the original
awkward example), \citet{Dreyer:jfp12} (DNB) formally capture the special status
of evaluation contexts using Kripke worlds with public and private future world
relations. Essentially, they allow relatedness of evaluation contexts to be
monotone with respect to a weaker future world relation (public) than
relatedness of values. This formalises the idea that it is safe to make
temporary internal state modifications (private world transitions, that
invalidate the continuation, but not other values) while a term is performing
internal steps, as long as it returns to a stable state (i.e. transitions to a
public future world of the original) before returning. We generalise this idea
to reason about local capabilities: validity of local capabilities is allowed to
be monotone with respect to a weaker future-world relation than other values,
which we can exploit to distinguish between state changes that are always valid
(public future worlds) and changes that are only valid if we clear local
capabilities (private future worlds). Our future world relations are similar to
\citet{Dreyer:jfp12}'s (for example, our proof of the awkward example uses the
exact same state transition system), but they turn up in an entirely different
place in the logical relation: rather than using public future worlds for the
special syntactic category of evaluation contexts, they are used in the value
relation depending on the locality of the capability at hand. Additionally, our
worlds are more complex because, to allow local memory capabilities and
write-local capabilities, they can contain (revokable) temporary regions that
are only monotonous w.r.t. public future worlds, while DNB's worlds are entirely
permanent.

\emph{Local capabilities in high-level languages} We point out that
local capabilities are quite similar to a feature proposed for the high-level
language Scala: \citet{osvald_gentrification_2016}'s second-class or local
values. They are a kind of values that can be provided to other code for
immediate use without allowing it to be stored in a closure or reference for
later use. We believe reasoning about such values will require techniques similar 
to what we provide for local capabilities.

\section{Related work}
\label{sec:related-work}

Finally, we summarise how our work relates to previous work. We do not
repeat the work we discussed in Section~\ref{sec:discussion}.

Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to
\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
of previous work. The capability machine formalized in
Section~\ref{sec:capab-mach-with} is a simple but representative
model, modeled mainly after the
M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers
resemble the M-Machine's) and
CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
memory and local capabilities resemble CHERI's). The latter is a
recent and relatively mature capability machine, which combines
capabilities with a virtual memory approach, in the interest of
backwards compatibility and gradual adoption. As discussed, our local
capabilities can cross module boundaries, 
contrary to what is enforced by CHERI's default CCall implementation.

Plenty of other papers\lau{If there are plenty, then I guess we should cite more
  than one?} enforce well-bracketed control flow at a low level, but most are
restricted to preventing particular types of attacks and enforce only partial
correctness of control flow. This includes particularly the line of work on
\emph{control-flow integrity}~\citep{abadi_control-flow_2005}. Those use a quite
different attacker model than us: they assume an attacker that is not able to
execute code, but can overwrite arbitrary data at any time during execution (to
model buffer overflows). By checking the address of every indirect jump and
using memory access control to prevent overwriting code, this work enforces what
they call control-flow integrity, formalized as the property that every jump
will follow a legal path in the control-flow graph. As discussed in
Section~\ref{sec:discussion}, such a property ignores temporal properties and
seems hard to use for reasoning.

More closely related to our work are papers that use a trusted stack manager and
some form of memory isolation to enforce control-flow correctness as part of a
secure compilation
result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
differs from theirs in that we use a different form of low-level security
primitive (a capability machine with local capabilities rather than a machine
with a primitive notion of compartments) and we do not use a trusted stack
manager, but a decentralized calling convention based on local
capabilities. Also, both prove a secure compilation result from a high-level
language, which clearly implies a general form of control-flow correctness,
while we define a logical relation that can be used to reason about
specific programs that rely on well-bracketed control flow.

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

El-Korashy also defined a formal model of a capability machine, namely CHERI,
and uses it to prove a compartmentalization
result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
correctness). He also adapts control-flow integrity (see above) to the machine
and shows soundness, seemingly without relying on capabilities.

%\section{Conclusion}
%\label{sec:conclusion}

\begin{comment}
\section*{Acknowledgements}
\label{sec:acknowledgements}

This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).
Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
\end{comment}

%\bibliographystyle{IEEEtran}
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,references}

\end{document}
\grid