### 2017-12-13 presentations/illustrations/local-prevention.tex

\begin{frame}
  \frametitle{Local Stack Capabilities prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw node[right=0cm of sp1] { stack pointer };

    % stores stack pointer in heap
    \begin{scope}
      \fill[gray!20] (14,6) rectangle (18,13);
      \draw (14,6) -- (14,13);
      \draw (18,6) -- (18,13);
    \end{scope}
    \draw (16,14) node {heap memory};
    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
\end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local caps: Stack clearing prevents Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker stores copy of stack pointer high in the stack
    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.-.(1)>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % trusted code clears the stack
    \begin{onlyenv}<+->
      \foreach \x in {9,9.5,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}
    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
    (13,13.5) -- (10,14.5);

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % attacker returns
    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
    \begin{onlyenv}<.>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % trusted code clears stack again
    \begin{onlyenv}<+->
      \foreach \x in {7,7.5,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}
    % then, the trusted code can return
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}

  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities prevent Attack 3}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    \begin{scope}
      \fill[gray!50] (13,0) rectangle (18,13);
      \draw (13,0) -- (13,13);
      \draw (18,0) -- (18,13);
      \draw (15.5,14) node {heap memory};
    \end{scope}

    \begin{onlyenv}<+->
      \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
      \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
    \end{onlyenv}

% Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
    \begin{onlyenv}<+->
      \begin{scope}
        \clip (6.1,2.5) rectangle (8.1,5.5);
        \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
      \end{scope}
      \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
      \draw node[right=0cm of ret1] { ret-ptr };
    \end{onlyenv}

    \begin{onlyenv}<+->
      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
      \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
      \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
    \end{onlyenv}

    \begin{onlyenv}<+-.(2)>
      \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
    \end{onlyenv}

    \begin{onlyenv}<+>
      \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
      \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
    \end{onlyenv}

    \begin{onlyenv}<+->
      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \end{onlyenv}
    \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);

    \begin{onlyenv}<+->
      \begin{scope}
        \clip (6.1,5.5) rectangle (8.1,8.5);
        \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
      \end{scope}
      \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
      \draw node[right=0cm of ret2] { ret-ptr (1) };
    \end{onlyenv}
    \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);

    \begin{onlyenv}<+->
      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
      \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
      \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};

    \end{onlyenv}

    \begin{onlyenv}<+->
      \foreach \x in {11.5,12,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}

    \begin{onlyenv}<+->
      \begin{scope}
        \clip (6.1,8.5) rectangle (8.1,11.5);
        \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
      \end{scope}
      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
      \draw node[right=0cm of ret3] { ret-ptr (2) };
    \end{onlyenv}

    \begin{onlyenv}<+->
      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
      \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
      \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
      \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
      \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
    \end{onlyenv}

    \begin{onlyenv}<+->
      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
    \end{onlyenv}
\end{tikzpicture}
\end{frame}
%%% Local Variables:
%%% TeX-master: "presentation-illustrations"
%%% End: