### 2017-12-13 presentations/illustrations/presentation-illustrations.tex

\documentclass{beamer}
\input{preamble}
\begin{document}

\input{traditional}

\input{attacks}

\input{local-prevention}

\input{linear-prevention}

\end{document}