### 2017-12-19 presentations/illustrations/presentation-illustrations.tex
Presentation illustrations: move text slide to main file.

\documentclass{beamer}
\input{preamble}
\begin{document}

\input{traditional}
\input{attacks}
\input{local-prevention}
\input{linear-prevention}

\begin{frame}
  \frametitle{Stack Capabilities Problems}
  \begin{itemize}
  \item Stack capabilities are not a panacea
  \item Stack pointers must not leave stack frame (through memory or stack)
  \item Similar problems for return pointers (omitted)
  \item Solution: local capabilities?
    \begin{itemize}
    \item Make stack and return capabilities local
    \item Cannot leave registers (except on stack) (\sout{Attack 1})
    \item Stack clearing on boundary crossings (\sout{Attack 2})
    \end{itemize}
  \end{itemize}
\end{frame}

\end{document}