### 2017-12-19 presentations/illustrations/presentation-illustrations.tex
used macro for temp registers in a couple of places

\documentclass{beamer}
\input{preamble}
\begin{document}

\input{semantics}

% \input{traditional}
% \input{attacks}
% \input{local-prevention}
% \input{linear-prevention}

\end{document}