### 2017-12-19 presentations/prisc18-01-2017-linear-caps/presentation.tex
Presentation for PriSC.

\documentclass{beamer}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}

\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}

\title{Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1}Aarhus University \qquad \inst{2}KU Leuven}
\date{PriSC 2018, Los~Angeles}

\input{../illustrations/preamble}
\begin{document}
\frame{\titlepage}

\begin{frame}
  \frametitle{A realistic secure compiler?}

  \vspace{-1cm}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % \draw<1-6> (-5,0) node {};
    % \draw (5,0) node {};
    % \draw (0,-14) node {};
    \draw (0,0) node[draw=black] (ir) {...};
    \draw node[draw=black,rounded corners,above=of ir] (source) {Source Language};
    \draw (source) edge[bend left, ->] (ir);
    \begin{onlyenv}<1-3>
      \draw node[draw=black,rounded corners,below=of ir] (target) {Target Language};
      \draw (ir) edge[bend left, ->] (target) ;
    \end{onlyenv}
    \begin{onlyenv}<2>
      \draw node[right=1cm of ir, line width=1mm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{onlyenv}
    \draw<3> node[right=1cm of target, line width=1mm, font=\fontfamily{ppl}\fontsize{1.5cm}{2cm}\selectfont] {?};
    \begin{visibleenv}<4-5>
      \draw (200:5) node[draw=black,rounded corners, text width=3.5cm] (commhar) {Commodity Hardware\\(Virtual Memory)};
      \draw (ir) edge[bend right=10, ->] (commhar) ;
      \draw (230:5) node[draw=black,rounded corners, text width=3.5cm] (wasm) {WebAssembly};
      \draw (ir) edge[bend left, ->] (wasm) ;
    \end{visibleenv}
    \begin{visibleenv}<4-6>
      \draw (280:5.8) node[draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
      \draw (ir) edge[bend left=20, ->] (dtal) ;
      \draw (335:8) node[draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
      \draw (ir) edge[bend right=20, ->] (mph) ;
      \draw (350:7) node[draw=black,rounded corners, text width=3.5cm] (capmach) {Capability Machines};
      \draw (ir) edge[bend left=10, ->] (capmach) ;
    \end{visibleenv}
    \begin{visibleenv}<5>
      \draw node[below left=0cm of commhar, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
      \draw node[below left=0cm of wasm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{visibleenv}
    \begin{visibleenv}<7>
      \draw node[draw=black,rounded corners,below=of ir] (capmach2) {Capability machines};
      \draw (ir) edge[bend left, ->] (capmach2) ;
    \end{visibleenv}
    \begin{visibleenv}<8->
      \draw node[draw=black,below=of ir] (ir2) {...};
      \draw (ir) edge[bend left, ->] (ir2) ;
      \draw node[draw=black,rounded corners,below=of ir2] (capmach3) {Capability machines};
      \draw (ir2) edge[bend left, ->] (capmach3) ;
    \end{visibleenv}
    \begin{visibleenv}<9>
      \draw node[rectangle,fit=(ir2) (capmach3),draw=red] (laststeps) {};
      \draw node[above right=0cm of laststeps,cloud,red] {Let's take a look!};
    \end{visibleenv}
  \end{tikzpicture}
  \only<5>{Fine-grained encapsulation?}
  \only<8>{Well-bracketed control flow, stack encapsulation?}
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

% \AtBeginSection[]
% {
%  \begin{frame}<beamer>
%  \frametitle{Outline}
%  \tableofcontents[currentsection]
%  \end{frame}
% }
\section{Stack and return pointers}
\input{../illustrations/traditional}

\begin{frame}<1>[label=sarps]
  \frametitle{Stack and return pointer security}
  \begin{itemize}
  \item Desirable properties:
    \begin{itemize}
    \item Well-bracketed control flow
    \item Stack frame encapsulation
    \end{itemize}
  \item How to enforce?
    \begin{itemize}
    \item Capabilities?
      \begin{itemize}
      \item<2-> Not enough: two attacks!
      \end{itemize}
    \item<3-> CHERI Local capabilities?
      \begin{itemize}
      \item Cannot leave registers
      \item<4-> Works, but... [Skorstengaard et al., ESOP 2018]
      \end{itemize}
    \item<5-> CHERI Linear capabilities?
      \begin{itemize}
      \item Non-copyable capabilities
      \item<6> Works perfectly?
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Stack and return capabilities}
\input{../illustrations/capabilities}

\section{Capabilities are not enough}
\againframe<2>{sarps}
\input{../illustrations/attacks}

\section{Local stack and return capabilities}
\againframe<3>{sarps}
\input{../illustrations/local-prevention}

\againframe<4>{sarps}
\section{Linear stack and return capabilities}
\againframe<5>{sarps}
\input{../illustrations/linear-prevention}
\againframe<6>{sarps}

\begin{frame}
  \frametitle{Formally?}

  \begin{itemize}
  \item<2-> Formulate correctness?
    \begin{itemize}
    % \item<3-> (CFI is too weak (right flows \emph{at the right time}))
    \item<3-> Idea: well-bracketed version of target assembly
    \item<3-> Compiler replaces call/returns with calling convention
    \item<3-> Fully abstract compilation
    \end{itemize}
  \item Prove correctness?
    \begin{itemize}
    \item<4-> Back-translation as embedding
    \item<4-> Cross-language step-indexed Kripke logical relation
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
  \item Enforcing well-bracketedness and stack frame encapsulation on a capability machine
  \item Local capabilities work [Skorstengaard et al, ESOP 2018]
  \item Linear capabilities work better!
    \begin{itemize}
    \item but don't exist... yet?
    \end{itemize}
  \item Correctness as fully abstract compilation from well-bracketed version of target language.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Thanks}
  \begin{center}
    \Large Questions?
  \end{center}
\end{frame}

\appendix
\begin{frame}
  \frametitle{Backup slide: The Awkward Example}
  \begin{align*}
    \tau &\isdef (unit \rightarrow unit) \rightarrow int\\
    e_1 &\isdef
          \left\{\begin{aligned}
            &\mathrm{let}~ x~ =~ \mathrm{ref}~ 0~ \mathrm{in}~\\
          &\lambda f\ldotp x \mathrel{:=} 0;~ f~\mathrm{unit}; x \mathrel{:=} 1; f~ \mathrm{unit}; !x
          \end{aligned}\right.\\
    e_2 &\isdef \lambda f\ldotp (f~\mathrm{unit}; f~\mathrm{unit}; 1)
  \end{align*}
\end{frame}

\end{document}