### 2018-04-15 presentations/esop18-04-2018/main.tex
Presentation notes

\documentclass{beamer}
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
  }

\usepackage{pgfpages}
\setbeameroption{show notes}
\setbeameroption{show notes on second screen=right}

\usepackage{listings}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\ttfamily,
  keywordstyle=\ttfamily\bfseries,
  morekeywords={let, ref, assert, in},
  columns=fullflexible,
  keepspaces=true,
  escapeinside={(*}{*)}
}

\input{../illustrations/preamble}
\input{loc_preamble}
\title{Reasoning About a Machine with Local Capabilities}
\subtitle{Provably Safe Stack and Return Pointer Management}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1} Aarhus University \and %
  \inst{2} imec-DistriNet, KU Leuven}

\date{ESOP, April 17, 2018}

\tikzset{
    gray on/.style={alt=#1{gray}{}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
}

\begin{document}
\maketitle

\begin{frame}[fragile]
  \frametitle{What Does This Program Do?}
  % Motivation
  \centering

  \begin{tabular}{c}
    \begin{lstlisting}
let x = ref 0 in
  (*$\lambda$*)f.(x := 0;
      f();
      x := 1;
      f();
      assert(!x == 1))
    \end{lstlisting}
  \end{tabular}

  \note<1>[item]{Consider program. Assuming a standard ML semantics we can say what it does.}
  \note<1>[item]{Bind x to freshly allocated reference in a closure that...}
  \note<1>[item]{takes callback f, sets x to 0, calls f, sets x to 1, calls f and finally asserts x points to 1.}
  \note<1>[item]{Note the assumption that when we call f, then we return to a specific program point. This is what we call well-bracketedness and we assume we have this in many programming languages.}
  \note<1>[item]{However, in order to execute this code, we need to compile it to assembly. }
  \note<1>[item]{How is well-bracketedness guaranteed? In particular, how is it guaranteed if f is a piece of code we do not trust (maybe handwritten assembly).}
  \note<2>[item]{\emph{We present a calling convention for capability machines that provide well-bracketedness and local state encapsulation as well as a logical relation that allows us to reason about such programs.}}
  \note<2>[item]{Let's first consider how stack pointers traditionally are handled.}

  % \note{Informally, what does this program do?}
  % \note{It allocates a reference to 0 and saves a closure to x. The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
  % \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
  % \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point. The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.} 
\end{frame}

\begin{frame}
  \frametitle{Traditional Stack Pointers}
  % \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
  \note<1-3>[item]{Simply put, a caller calls a function which}
  \note<1-3>[item]{pushes a new stack frame on the stack the callee uses for its execution. }
  \note<1-3>[item]{When the callee is done, then it returns to the caller by popping its stack frams.}
  \note<4->[item]{If \emph{callee (evil) assembly code with no intention to follow the CC}, then there are multiple ways for them to break things:}
  \note<5->[item]{Read or write directly from or to the caller's stack frame, breaking local-state encapsulation}
  \note<6->[item]{Skip the caller's stack frame and return to one further down breaking well-bracketedness.}
  \note<6->[item]{Clearly we need some kind of low-level enforcement mechanism.}

  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
    \end{scope}
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);

    % traditional:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);

    % traditional return:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call, attack:
    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };

        % read other stack frames
    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.-.(1)> (sp0) edge[->,out=235,in=0] (6,8.25);

    % read other stack frames

    \actadv[+]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);

    % break well-bracketedness
    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
      skip caller frame};
  \end{tikzpicture}
  % \note{We need something to enforce security properties on a low-level machine.}
\end{frame}

\begin{frame}
  \frametitle{Capability Machine}
  % short description of capability machines

  \begin{columns}
\begin{column}{0.5\textwidth}
  \begin{itemize}
  \item<1-> Low-level machine
    \note<1->[item]{Capability machines are low-level machines proposed in the systems community.}
    \note<1->[item]{For instance, the CHERI OS operates on one.}
    \note<1->[item]{Has all the instructions we expect, load, store, jmp, etc.}
    % \note{Has instructions as you would expect, load, store, jmp, etc.}
  \item<2-> Capabilities replace pointers
    \begin{itemize}
    \item<3-> Pointer
    \item<4-> Range of authority
    \item<5-> Kind of authority
      \note<5->[item]{Roughly two kinds of capabilities:}
      \begin{itemize}
      \item read, write, and execute
        \note<5->[item]{Memory capabilities, allows you to do all the standard memory operations.}
      \item enter
        \note<5->[item]{Provides encapsulation mechanism which allows seperation of security domains.}
        \note<5->[item]{Can not be used for anything but jump, when jumped to becomes read/execute.}
      \end{itemize}
    \end{itemize}
    \item<6-> Capability manipulation instructions
    \item<7-> Authority checked dynamically
    \end{itemize}
\end{column}

\begin{column}{0.5\textwidth}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \draw (1,9) node {Memory};
    \draw (-3.5,0) node {\phantom{.}};
    \scope
      \clip (-.1,-.1) rectangle (6.1,8.1);
      \fill[fill=white] (0,0) rectangle (6,8);
      \draw (0,0) -- (0,8);
      \draw (6,0) -- (6,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    \draw<3>  (-3,1.5) edge[black,->,in=180] (0,2.75);

    \draw<4-> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
    (-0.25,1) -- (-0.25,4.5) node[draw=black] (sp1) [black,midway,xshift=-12pt] {};
    \draw<4>  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize\phantom{\textsc{rw}}} (sp1);
    \draw<5->  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize \textsc{rw}} (sp1);
  \end{tikzpicture}
\end{column}
\end{columns}
 % \note{This particular setup gives a very fine-grained memory control.}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 1}
%  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
  \note[item]<1->{Let's see how this changes things: Now the \emph{untrusted code cannot immediately read from the caller stack frame}, because the stack capability does not have authority over that part of memory.}
  \note[item]<1->{There is nothing that prevents the \emph{untrusted code from storing the stk ptr on the heap.}}
  \note[item]<3->{Upon return the \emph{callee regains its stack capability} which has authority over the callee stack frame and everything above. }
  \note[item]<3->{The \emph{caller pushes some important things} on the stack and \emph{calls the untrusted code again}. With a smaller stack pointer.}
  \note[item]<4->{The stack pointer the caller gives the untrusted code cannot be used to access the callee stack frame, but because the untrusted code stored the old stack pointer, it now has access to part of the callee's stack frame.}
  \note[item]<4->{Again breaking local state encapsulation.}
  \note[item]<4->{Need a way to make sure stack pointer is not stored for later use.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \begin{onlyenv}<+-+(2)>
      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node[above right=1cm of sp1, xshift=-0.5cm] (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
    \end{onlyenv}

%    \note{Evil callee stores stack pointer on the heap.}
    % stores stack pointer in heap
    \begin{onlyenv}<+->
      \begin{scope}
        \fill[gray!20] (14,6) rectangle (18,13);
        \draw (14,6) -- (14,13);
        \draw (18,6) -- (18,13);
      \end{scope}
      \draw (16,14) node {heap memory};
      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \end{onlyenv}
    \draw<.> (14,8.25) edge[->,red,thick,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (14,8.25) edge[->,red,thick,bend left] node[left] {\footnotesize \textsc{rw}}(spold);
    \end{onlyenv}

%    \note{Evil callee returns to the caller.}
    % attacker returns
    \begin{onlyenv}<+>
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below right] {\footnotesize \textsc{rw}} (sp1);
      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
    \end{onlyenv}

%    \note{caller uses a bit more of the stack and calls the evil callee again.}
%    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
    % attacker gets called again + uses old stack pointer.
    \begin{onlyenv}<+>
      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,14) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local Capabilities}
  % Short description of local capabilities
  CHERI inspired
  \begin{itemize}
  \item Capabilities tagged with locality (local or global)   %\note{We call non-local capabilities global}
  \item New \textit{write-local} permission
  \item Local capabilities can only be stored by capabilities with \textit{write-local} permission
  \end{itemize}
  \note[item]<1->{To revoke a capability, we need to find it in memory which means we need access + need to search the entire memory.}
  \note[item]<1->{Restricted where local capabilities can be stored. restricts where we need to look for a capability.}
  \note[item]<1->{We define a calling convention. In order to prevent attack 1, we do the follwoing.}
  \onslide<2->{Calling convention}
  \begin{itemize}[<2->]
  \item Stack capability is local with permission read, write-local, and execute.
    \note[item]<2->{Local stack capability cannot be stored on the heap. We need to be able to store old stack pointers somewhere, traditionally stack. }
  \item No global write-local capabilities.
    \note[item]<2->{Global write-local capabilities would undermine the entire idea as it would allow local capabilities to be stored indirectly.}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities Prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \note{In the attack from before, when the attacker \emph{attempts to store the stack capability} on the heap, then the \emph{machine checks} that we have the correct authority to perform the operation. \emph{Assuming we only have global capabilities} for the heap, it cannot have write-local authority, due to the assumption on the previous slide, so we try to store the stack capability through a capability that does not have write-local authority, so it fails.}

    % recurrent parts
    \stdstackstart

    % attacker...
    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw (11,14) node (a) { stack pointer };
    \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);

    % stores stack pointer in heap
    \begin{scope}
      \fill[gray!20] (14,6) rectangle (18,13);
      \draw (14,6) -- (14,13);
      \draw (18,6) -- (18,13);
    \end{scope}
    \draw (16,14) node {heap memory};
    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw (14,8.25) edge[->,red,thick,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
\end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \note[item]{While this prevents attack 1, we are not quite safe done.}
    % recurrent parts
    \stdstackstart
    \note[item]{Trusted caller calls untrusted code.}
    \note[item]<2->{untrusted code stores the stack pointer on the stack.}
    \note[item]<2->{stack pointer local, but stack pointer has write local permission, so no problem.}
    \note[item]<3->{untrusted code calls some trusted code with a callback.}
    \note[item]<5->{trusted code runs for a bit pushes some local data to the stack and calls the callback.}
    \note[item]<6->{The stack pointer is still on the stack allowing the untrusted code to read write to the stack frame of the trusted code.}
    % first step: capabilities
    \begin{onlyenv}<+>
      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
    \end{onlyenv}

    % attacker stores copy of stack pointer high in the stack
    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<.-.(2)>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);        
    \end{onlyenv}
    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.> (6,13.25) edge[red,thick,->,in=45] node[above right,yshift=0.1cm,xshift=-0.2cm] {\footnotesize local,\textsc{rwlx}}(sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (6,13.25) edge[red,thick,->,in=45] node[sloped, above] {\footnotesize local,\textsc{rwlx}}(spold);
    \end{onlyenv}

    % attacker calls trusted code again
%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
        \draw (11,12.5) node (a) { stack pointer };
        \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \end{onlyenv}

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.-.(1)>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};

      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \end{onlyenv}
    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
%  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
\end{frame}

\begin{frame}
  \frametitle{Calling Convention (Continued)}
  \note[item]{Stack is basically the only place we can store local capabiliites.}
  \note[item]{Make sure that untrusted code don't ``sneak'' capabilities between calls on the stack}
  \note[item]{Clear stack and argument registers}
  $\dots$
  \begin{itemize}
  \item Clear stack and non-argument registers before invoking untrusted code.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Stack Clearing Prevents Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \note[item]{Let's see that the addition to the CC prevents attack 2.}
    \note[item]{The untrusted code has been called. It calls the well-behaved code.}
    \note[item]{The well-behaved code does its thing, but this time it clears the stack overwritting the old stack pointer the untrusted code had saved for later.}
    \note[item]{The untrusted code starts running, but it does not have an old stack pointer available only the one given to them by the well-behaved code}
    % recurrent parts
    \stdstackstart
    % attacker stores copy of stack pointer high in the stack
    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.-.(1)>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
    \end{onlyenv}

    % trusted code clears the stack
    \begin{onlyenv}<+->
      \foreach \x in {9,9.5,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}
    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
    (13,13.5) -- (10,14.5);

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw (11,12.5) node (a) { stack pointer };
      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);      
    \end{onlyenv}

    % % attacker returns
    % \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
    % \begin{onlyenv}<.>
    %   \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    %   \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    %     (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    %   \draw node[right=0cm of sp1] { stack pointer };
    % \end{onlyenv}

    % % trusted code clears stack again
    % \begin{onlyenv}<+->
    %   \foreach \x in {7,7.5,...,14}
    %   {
    %     \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
    %   };
    % \end{onlyenv}
    % % then, the trusted code can return
    % \begin{onlyenv}<+>
    %   \actadv{(0,4)}{(6,7)}{callee stack frame}
    % \end{onlyenv}

  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{(Full) Calling Convention}  
  % and stack management
  \note[item]{The calling convention contains a bit more, but all of it is motivated by some attack.}
  \note[item]{I won't motivate the rest here, but I wanted to show you that it does not take many more precautions.}
  \begin{itemize}
  \item Initially:
    \begin{itemize}
    \item \textit{Stack capability local capability with read, write-local, and execute authority.}
    \item \textit{No global write-local capabilities on the machine.}
    \end{itemize}
  \item Prior to returning to untrusted code:
    \begin{itemize}
    \item \textit{Clear the stack.}
    \item Clear non-return registers.
    \end{itemize}
  \item Prior to calls to untrusted code:
    \begin{itemize}
    \item Push activation record to the stack and create enter-capability.
    \item Restrict the stack pointer to the unused part and clear that part.
    \item Clear non-argument registers.
    \end{itemize}
  \item Only invoke global call-backs.
  \item When invoked by untrusted code
    \begin{itemize}
    \item Make sure the stack pointer has read, write-local and execute authority.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Formalizing the Guarantees of a Capability Machine}
  \note[item]{How can we be sure the calling convention works?}
  \note[item]{Specifically, if we have a \emph{program that interacts with untrusted code using the calling convention}, how do we formally show correctness of the program.}
  \note[item]{We need a formal statement about the guarantees provided by the capabilities including the specific guarantees for local capabilities.}
  \note[item]{Traditionally syntactic very syntactic (e.g. reference graph), does not take into account what the program does with its capabilities.}
  \note[item]{\emph{We} have defined a logical relation which also give us a statement about the guarantees provided by the capability machine.}
  \begin{itemize}
  \item How can we be sure the calling convention works?
  \item<2-> Unary step-indexed Kripke logical relation over recursive worlds
    \begin{itemize}
    \item Statement of guarantees provided by the capability machine
    \end{itemize}
  \end{itemize}
  \note[item]<2->{Calling convention main application, but it is general}

  \note[item]<2->{In the following: give some intuition about what a LR looks like for a capability machine}

  % \note{correctness here could be assert not violated}

  % \note{mention better than previous. Define the guarantees}
\end{frame}

\begin{frame}
  \frametitle{Worlds, Safe Values, and Step-Indexing}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
      \note[item]<1->{Compared to standard assembly language, capabilities executing code has access to represent bound.}
    \item<1-> Capabilities represent bounds on executing code
      \note[item]<1->{That is, the capabilities the executing code has access to.}
    \item<2-> World, $W$      
      \begin{itemize}
      \item Collection of invariants
      \end{itemize}
    \item<4-> Predicate for safe values w.r.t world, $\stdvr(W)$
      \begin{itemize}[<8->]
      \item Recursively defined
      \end{itemize}
    \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \scope
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[fill=white] (0,0) rectangle (6,15);
      \draw (0,0) -- (0,15);
      \draw (6,0) -- (6,15);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    \foreach \x in {0.5,1,...,14}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    % One capability
    \draw<1-6> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
    \draw<1-4>  (6,7.25) edge[black,->,in=45] (sp1);

    % Invariant colored
    \fill<2->[red,opacity=.7] (0,1) rectangle (6,6);
    \fill<2->[blue,opacity=.7] (0,7) rectangle (6,10);

    \fill<3-6>[black, fill opacity=0, text opacity=0.7] (0,4) rectangle (6,4.5) node[pos=.5] {\footnotesize 42};

    \draw<5> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{r}} (sp1)   ;
    \draw<6> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{rw}} (sp1)   ;
    \draw<6>[very thick,color=red] (6.5,6.5) -- (8.5,4.5)
    (6.5,4.5) -- (8.5,6.5);

    \draw<7-> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
    \path<7->  (6,4.25) edge[black,out=30,in=135] node[above right]  {\footnotesize \textsc{r}} (8,5) 
              (8,5) edge[black,->,out=-45,in=45](sp1);

  \end{tikzpicture}

\end{column}
\end{columns}
  % % Do this as a series of figures?

  % % Figure with memory and one capability to make the bound explicit
  % \note{compared to normal assembly, capabilities represent bounds on executing block of code.}

  % \note{we have no observable I/O, so the authority bounds we consider are related to memory. However, more fine-grained/detailed than read/write-authority.}

  % % Color the memory to illustrate different regions of the world
  % \note{a piece of code can be bound by arbitrary memory invariants which we define in a \emph{world}.}

  % % Write down a static invariant (some address contains 42, maybe write a legend on the left side of slide.).
  % \note{essentially, a world is a collection of invariants and safety of words defined with respect to a world.}

  % \note{Define a set of words that are safe w.r.t. world W V(W) in P(Word)}

  % \note{Whether a capability is safe depends on the authority it carries}

  % % Add a read capability to the address that contains 42, safe
  % % Add read/write capability to the address that contains 42, not safe - can be used to violate the invariant.
  % \note{Example, w.r.t. a world with the invariant that an address contains constant. Safe for read capability, not write as write capability can break this invariant.}

  % % Add a read capability that refers to itself
  % \note{safety for a read capability is the case if the read capability only gives access to safe capabilities.}

  % \note{What if that part of memory contains a read capability for the same part of memory? Cyclic definition. Solved by step-indexing.}

  % \note{related to similar issue with languages with recursive types or higher-order ML-list references.}

  % \note{Solved by step-indexing - safety up to a certain number of interaction steps.}
\end{frame}

\begin{frame}
  \frametitle{Future Worlds and Invariants, and Recursive Worlds}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \scope
      \clip (-.1,-.1) rectangle (6.1,8.1);
      \fill[fill=white] (0,0) rectangle (6,8);
      \draw (0,0) -- (0,8);
      \draw (6,0) -- (6,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    \fill[red,opacity=.7] (0,7.5) rectangle (6,6);
    \fill[blue,opacity=.7] (0,6) rectangle (6,4);

    \begin{onlyenv}<2->
    \scope
      \clip (11.9,-.1) rectangle (18.1,8.1);
      \fill[fill=white] (12,0) rectangle (18,8);
      \draw (12,0) -- (12,8);
      \draw (18,0) -- (18,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (12,\x) rectangle (18,\x+.5) node[pos=.5,color=teal] {};
    };

    \fill[red,opacity=.7] (12,7.5) rectangle (18,6);
    \fill[blue,opacity=.7] (12,6) rectangle (18,4);

    \draw [decorate,decoration={brace,amplitude=3pt,raise=2pt},yshift=0pt]
    (11.75,1) -- (11.75,4) node[draw=black] (sp1) [black,midway,xshift=-12pt] {}; % ,shape=coordinate,xshift=1pt
    \draw  (10,0.5) edge[black,->,in=225] (sp1);

    \draw (7,7.5) edge [black,thick, ->] node[above] {\footnotesize Memory allocated} (11,7.5);
  \end{onlyenv}
  \begin{onlyenv}<3->
    \fill[green,opacity=.7] (12,4) rectangle (18,1);

    \draw (9,4) node {\huge $\sqsubseteq$};
  \end{onlyenv}
    %Invariant transition systems
    \begin{onlyenv}<4->
    \node[circle,draw=black, fill=white] (a) at (3,6.75) {};
    \draw[>=stealth] (2,6.75) edge[black,->] (a);

    \node[circle,draw=black, fill=white] (b) at (2,5) {};
    \draw[>=stealth] (1,5) edge[black,->] (b);
    \node[circle,draw=black, fill=white] (c) at (4,5) {};
    \draw[>=stealth] (b) edge[black,->, bend left] (c);
    \draw[>=stealth] (c) edge[black,->, bend left] (b);

    \node[circle,draw=black, fill=white] (a2) at (15,6.75) {};
    \draw[>=stealth] (14,6.75) edge[black,->] (a2);

    \node[circle,draw=black, fill=white] (b2) at (14,5) {};
    \draw[>=stealth] (13,5) edge[black,->] (b2);
    \node[circle,draw=black, fill=white] (c2) at (16,5) {};
    \draw[>=stealth] (b2) edge[black,->, bend left] (c2);
    \draw[>=stealth] (c2) edge[black,->, bend left] (b2);

    \node[circle,draw=black, fill=white] (d) at (14,2.5) {};
    \draw[>=stealth] (13,2.5) edge[black,->] (d);
    \node[circle,draw=black, fill=white] (e) at (16,3.25) {};
    \draw[>=stealth] (d) edge[black,->, bend left] (e);
    \draw[>=stealth] (e) edge[black,->, bend left] (d);
    \node[circle,draw=black, fill=white] (f) at (16,1.75) {};
    \draw[>=stealth] (e) edge[black,->, bend left] (f);
  \end{onlyenv}
\end{tikzpicture}

  \begin{overprint}
\onslide<1-4>
  \begin{itemize}
  \item<1-> Memory evolves over time
  \item<3-> Add invariants in future worlds
  \item<4-> Invariants as state machines
  \end{itemize}
  \onslide<5-6>
  \begin{itemize}
  \item<5-> Each state contains a predicate of accepted memory segments
  \item<6-> World indexed
  \end{itemize}
  \[
    H \; :  \only<5>{\phantom{\; \Worlds \fun {}}}\only<6>{\; \Worlds \fun} \mathrm{Pred}(\mathrm{MemSeg})
  \]
\end{overprint}
  \note{Memory changes over time, for instance new memory may be allocated.}
  \note{Allow worlds to evolve. New invariants can be added to handle freshly allocated memory.}
  \note{Safety of words monotone w.r.t. worlds (which makes it into a Kripke Logical relation).}
\end{frame}

% \begin{frame}
%   \frametitle{Invariants and Recursive Worlds}
%   \note{The purpose of a piece of memory may change over time. Invariants need to be more like protocols.}
%   \note{Take a closer look at invariants.}
%   \note{Each invariant is a state machine to model changes of memory invariants for a piece of code over time.}

%   \note{An invariant contains a mapping from states to predicate of memsegs...

%  for pieces of memory it accepts. Predicate may be world dependent (as the validity may depend on the remainder of the memory, which for instance is the case if it contains a capability(?)) or we might just want to say a specific address contains some safe value, but safety is world dependent. This means that the predicate must be world indexed.}
% \end{frame}

\begin{frame}[fragile]
  \frametitle{Local Capabilities}
  \texttt{f} is unknown code and \texttt{c} is a capability.
  \begin{lstlisting}
    f(c);
    f(1)
  \end{lstlisting}
  \begin{itemize}
  \item<2-> \texttt{c} global $\Rightarrow$ available in second invocation of \texttt{f}

    \note{when c global, second invocation in a world where c is safe}
  \item<3-> \texttt{c} local $\Rightarrow$ not available in second invocation of \texttt{f}
    \note{when c local, second invocation in a world where c is not (necessarily) safe}
    \note{c essentially revoked, so the invariants it relies on need not hold.}
  \end{itemize}

  \note{Two future world relations, one for all capabilities and one for non-local capabilities}
  \note{If c global, then second invocation must happen in public future world, so c valid.}
  \note{If c local, then second invocation may happen in a private future world.}
  \begin{onlyenv}<4->
  \begin{lemma}[Double monotonicity of value relation]
    \begin{itemize}
    \item If $\npair{w} \in \stdvr(W)$ and $W' \futurewk W$ then $\npair{w} \in
      \stdvr(W')$.
    \item If $\npair{w} \in \stdvr(W)$ and $W' \futurestr W$ and $w$ is
      not a local capability, then $\npair{w} \in \stdvr(W')$.
    \end{itemize}
  \end{lemma}
\end{onlyenv}
  \note{How does local/global capabilities affect all this.}
  \note{If we hand a global capability to untrusted code, then it may be stored in memory, so we will only be able to reinvoke that code if we can guarantee that those values are still valid. }
  \note{Formally, the worlds contain the invariants that the global capability depend on and reinvocation is only possible in future worlds where these invariants are respected.}
  \note{Local capabilities on the other provides a means to revoke capabilities. If we invoke untrusted code and give them a local capability, then they have no way to store it aside from the register file (and the stack), so when they return we can be sure that the local capability resides nowhere but the register file and the stack.}
  \note{If we clear the register file and stack for old local capabilities before we invoke the untrusted code again, then we have essentially revoked those local capabilities and we can safely invoke the untrusted code again with the assumption that it does not have access to the local capability from the first call.}
  \note{Formally, we allow this call to happen in a \emph{private future} world in which all global capabilities remain safe, but local capabilities don't.}
  \note{We also have a \emph{public future} world relation in which \emph{all} capabilities remain safe.}
  \note{Concretely, worlds may contain temporary regions on which local capabilities may depend. These regions persist in public future worlds which allows the local capabilities to rely on them. On the other hand, in private future worlds temporary regions may be revoked which is why local capabilities may not be safe in private future worlds.}

  \note{include comment about priv/pub future worlds + interplay with monotonicity?}
\end{frame}

\begin{frame}
  \frametitle{Fundamental Theorem of Logical Relations}
    \begin{itemize}
  \item General statement about the guarantees provided by the capability machine.
  \item Intuitively: any program is safe as long as it only has access to safe values.
  \end{itemize}
  \begin{theorem}[Fundamental theorem (simplified)]
      If 
      \[
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \]
      then
      \[
        \npair{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W)
      \]
  \end{theorem}

  % Comments/present high-level perspective
  \note{readCond is the assumption that every thing in the interval [a,e] is safe to read.}
  \note{$\stder$ is safe to execute relation. That is, it will respect all the memory invariants.}
  \note{That is take an arbitrary capability. If it only has access to safe capabilities then it will preserve the invariants of the world.}
  \note{Remeber, dynamic checks = failing is concidered secure} 
\end{frame}

\begin{frame}[fragile]
  \frametitle{``Awkward Example''}
    \begin{tabular}{c}
    \begin{lstlisting}
let x = ref 0 in
  (*$\lambda$*)f.(x := 0;
      f();
      x := 1;
      f();
      assert(!x == 1))
    \end{lstlisting}
  \end{tabular}
\note{example known from the literature}
\note{even just in ML difficult as f can be the closure.}
\note{the assert can fail if the calls are not well-bracketed!}  
\note{the local state is difficult to handle as the closure and the context needs to be able to update the invariant for x in different ways. (closure can switch between 0 and 1 as it pleases, but context can transition only from 0 to 1.)}
\note{relies heavily on well-bracketedness}
\note{we have made a faithful translation and proved correctness (i.e., the assertion never fails).}
\note{more semantic statement of guarantees allows us to do this.}
\end{frame}

% \begin{frame}
%   \frametitle{Contributions}
%   \begin{itemize}
%     \item Formalisation of a capability machine with local capabilities
%     \item Novel calling convention for enforcing control-flow correctness and local-state encapsulation
%     \item A calling convention utilizing local capabilities to ensure well-bracketed control-flow and local-state encapsulation
%     \item Step-indexed Kripke Logical Relation to reason about programs on said machine
%       \begin{itemize}
%         \item Single orthogonal closure
%         \item Variant of public/private future world relation to express nature of local capabilities
%       \end{itemize}
%       %new insights for low-level systems?
%     \item FTLR
%       \begin{itemize}
%       \item Semantic statement of guarantees offered by capability machine
%       \end{itemize}
%     \item Logical relation applied to a number of examples
%       \begin{itemize}
%       \item ``Awkward example''
%       \end{itemize}
%     \end{itemize}
% \end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
  \item Capability machines can guarantee properties of high-level languages.
  \item We developed a calling convention that ensures well-bracketedness and local-state encapsulation.
  \item We define a unary step-indexed Kripke logical relation over recursive worlds.
    \begin{itemize}
    \item Formal statement about guarantees provided by capability machine.
    \item Allows reasoning about programs on capability machine.
    \end{itemize}
  \item We apply it on the ``awkward example''.
  \end{itemize}
\end{frame}

\begin{frame}
  \centering
  Thank you!
\end{frame}

\end{document}