### 2019-01-10 presentations/popl19-01-2019/presentation.tex
More presentation work

\documentclass[aspectratio=169]{beamer}
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
  }

\usepackage{xcolor}
\input{preamble}

\usepackage{pgfpages}
\setbeameroption{hide notes}
%\setbeameroption{show notes on second screen=right}

\definecolor{myred}{RGB}{192,58,45}
\definecolor{mypurple}{RGB}{180,52,202}
\definecolor{myblue}{RGB}{51,51,179}
\definecolor{mylblue}{RGB}{52,178,204}
\definecolor{mygreen}{RGB}{66,192,103}

\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{calc}
\usepackage{graphicx}
\usepackage{dashrule}

\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape=true,
  literate={\$}{\$}{1},
  keywordstyle=\color{mylblue},
  tabsize=2
}

\lstdefinelanguage
   [x64]{Assembler}     % add a "x64" dialect of Assembler
   [x86masm]{Assembler} % based on the "x86masm" dialect
   % with these extra keywords:
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD, %
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS, %
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b, %
                  r10,r10d,r10w,r10b,r11,r11d,r11w,r11b, %
                  r12,r12d,r12w,r12b,r13,r13d,r13w,r13b, %
                  r14,r14d,r14w,r14b,r15,r15d,r15w,r15b}} % etc.

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Overview}
    \tableofcontents[currentsection]
  \end{frame}
}

%\input{../illustrations/preamble}
%\input{loc_preamble}
\title{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1} Aarhus University \and %
  \inst{2} Vrije Universiteit Brussel}

\date{POPL, January 19, 2019}

\begin{document}
\maketitle

\newsavebox{\assem}
\begin{lrbox}{\assem}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
main:
	.cfi_startproc
# BB#0:
	pushq	%rbp
.Ltmp0:
	.cfi\_def\_cfa\_offset 16
.Ltmp1:
	.cfi\_offset %rbp, -16
	movq	%rsp, %rbp
.Ltmp2:
	.cfi\_def\_cfa\_register %rbp
	subq	$\mbox{\textdollar}$16, %rsp
	movabsq	$\mbox{\textdollar}$.L.str, %rdi
	movl	$\mbox{\textdollar}$0, -4(%rbp)
	movb	$\mbox{\textdollar}$0, %al
	callq	printf
	xorl	%ecx, %ecx
	movl	%eax, -8(%rbp)
	movl	%ecx, %eax
	addq	$\mbox{\textdollar}$16, %rsp
	popq	%rbp
	retq
.Lfunc\_end0:
	.size	main, .Lfunc\_end0-main
	.cfi\_endproc
\end{lstlisting}
\end{lrbox}

\newsavebox{\hello}
\begin{lrbox}{\hello}
\begin{lstlisting}[basicstyle=\tiny\ttfamily, language=C, keywordstyle=\color{mylblue}]
#include <stdio.h>
int main()
{
  printf("Hello, World!");
  return 0;
}
\end{lstlisting}
\end{lrbox}

\begin{frame}
  \frametitle{Overview}
  \tableofcontents
\end{frame}

\section{\stktokens{}-paper in the big picture}
\begin{frame}[fragile]
  \frametitle{Abstractions all the way down}
  \centering
  \begin{tikzpicture}
    % \draw (9.5,0.25) rectangle (13.5,4.75);
    \draw (0,0) node {};
    \node<1->[inner sep=0pt, anchor=west] (hardware) at (9.5,2.5) {\includegraphics[angle=90,width=.285\textwidth]{hardware3-2}};

    \draw<2->[dashed] (9,-.5) -- (9,5.5);
%    \draw (4.75,n0.25) rectangle (8.75,4.75);
    \node<2->[inner sep=0pt, anchor=west] (assem) at (4.4,2.5) {\usebox{\assem}};

    \draw<3->[dashed] (3.9,-.5) -- (3.9,5.5);
%    \draw (0,0.25) rectangle (4,4.75);
    \node<3->[inner sep=0pt, anchor=west] (hello) at (0,2.5) {\usebox{\hello}};

    \draw<4> (.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center] node[above] {\phantom{secure}\\compilation} (4.3, 3.7);
    \draw<5-6> (.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center] node[above] {{\color{myred} secure}\\compilation} (4.3, 3.7);
    \node<6->[rectangle,rotate=-25,fill=white,rounded corners,draw] at (11.5,2.5) {{\color{myred}Capability machine}};

    \begin{onlyenv}<7->
      \draw (0.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center,color=white] node[above] {{\color{myred} secure}\\compilation} (4.3, 3.7);
      \draw node[rectangle,fill=white,rounded corners,draw] at (1.93,3.7) (ir1) {ir};
      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {ir'};
      \draw (0.8, 3.7) edge[-stealth,bend left,very thick] node[above] {} (ir1);
      \draw (ir1) edge[-stealth,bend left,very thick] node[above] {} (ir2);
      \draw (ir2) edge[-stealth,bend left,very thick] node[above] {} (4.3,3.7);
    \end{onlyenv}

    \draw<8-> node at (4.3,3.7) (p) { };
    \draw<8-> node[rectangle, fit=(ir2) (p),draw={rgb:red,192;green,58;blue,45}] (lastphase) {};
  \end{tikzpicture}

    \note[item]{Compbuter at the hardware level is a lot of physics going on very, very quickly.}
    \note[item]{Abstractions: instead of currents, bits.}
    \note[item]{Assembler - a lot more abstraction. We can make sense of it, but still not nice to work with.}
    \note[item]{High-level languages - enough abstractions to work with it and
      build incredible software systems.}
    \note[item]{In order to move between abstractions, we compile programs}
    \note[item]{Most compilers do not provably preserve the abstractions of
      high-level language.}
    \note[item]{This leads to security vulnerabilities}
    \note[item]{Important for compilation to preserve abstractions.}
    \note[item]{A compilation that preserves abstractions, even when linked with
      unknown assembly,is known as a \emph{secure} compilation}
    \note[item]{Little hope that it is possible to make a secure compilation to
      a traditional processor - they have no way to enforce the abstraction.}
    \note[item]{Necessary with hardware with enforcement mechanisms.}
    \note[item]{Our take is that this could be a capability machine}
    \note[item]{Capability machine explained briefly: Basically a processor as
      you know it except pointers are replaced with capabilities. Capabilities
      do not just point to memory, they also carry permissions that specify what
    operations they can be used for and on what part of the memory they can operate.}
    \note[item]{Naturally, this means that the assembly language must be
      capability aware.}
\end{frame}

% 	.type	.L.str,@object          # @.str
% 	.section	.rodata.str1.1,"aMS",@progbits,1
% .L.str:
% 	.asciz	"Hello, World!"
% 	.size	.L.str, 14

% #include <stdio.h>
% int main()
% {
%   printf("Hello, World!");
%   return 0;
% }

%\section{Paper Overview}
\begin{frame}
  \frametitle{Paper Contents Overview}
  \begin{description}
  \item[Formalization of CHERI-like capability machine with linear capabilities]
  \item[\stktokens{}] a calling convention that provably guarantees local-state
    encapsulation (LSE) and well-bracketed control-flow (WBCF)
  \item[Fully-abstract overlay semantics] a novel way to prove LSE and WBCF claims
  \end{description}
\end{frame}

\section{Defining well-bracketed control flow and local state  encapsulation}

\newcommand{\drawpc}[1]{

}

\begin{frame}[fragile]
  \frametitle{Defining well-bracketed control flow and local state encapsulation}
  \begin{columns}
    \begin{column}{0.5\textwidth}
\begin{lstlisting}[language=C, escapechar=!]
void a()       !\tikz[remember picture] \node[] (bt) {};! 
{
  !\tikz[remember picture] \node[] (pc6) {};!...
  !\tikz[remember picture] \node[] (pc7) {};!return;      
}              !\tikz[remember picture] \node[] (bb) {};!

void b()
{
  !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
  !\tikz[remember picture] \node[] (pc2) {};!a();
  !\tikz[remember picture] \node[] (pc3) {};!...
  !\tikz[remember picture] \node[] (pc4) {};!a();
  !\tikz[remember picture] \node[] (pc5) {};!return;   
}
\end{lstlisting}
      \begin{tikzpicture}[remember picture, overlay, text width=3cm]
%        \draw<1> node { };
        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=-1cm] (brace) [midway] {};
        \draw<2> node[right = -0cm of brace] {Function \texttt{a} cannot access variable \texttt{x}};

        \draw<3-> ($(pc1) + (-0.5,0)$) edge[-stealth] node {} (pc1);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.5\textwidth}
      \only<2>{Local-state encapsulation}
      \only<3->{Well-bracketed control flow}
    \end{column}
    % What is LSE and WBCF
    % Quick example?
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Desired properties of the WBCF and LSE definition}
  \begin{enumerate}
\item {\itshape Intuitive}%Should correspond to our intuition about WBCF and LSE
\item {\itshape Useful for reasoning}% we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}.
\item {\itshape Reusable in secure compiler chains}% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
\item {\itshape Arguably "complete"}% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
\end{enumerate}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Overlay Semantics}

\begin{tabular}{p{3.2cm} p{3.2cm}}
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
   !\tikz[remember picture] \node[] (srccodetl) {};!move  rtmp1 42
   !\tikz[remember picture] \node[] (calltl) {};!store rstk rtmp1
    cca   rstk -1
    geta  rtmp1 rstk
    cca   rretc 5      !\tikz[remember picture] \node[] (callbr) {};!
    move  rtmp1 pc
    cca   rtmp1 -20
\end{lstlisting}&
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
    load  rtmp1 rtmp1
    cca   rtmp1 -21
    cseal rretd rtmp1
    move  rretc pc
   !\tikz[remember picture] \node[] (retl) {};!xjmp  r1 r2 !\tikz[remember picture] \node[] (retr) {};!
    cseal rretc rtmp1
    move  rtmp1 0    !\tikz[remember picture] \node[] (srccodebr) {};!
\end{lstlisting}
\end{tabular}\tikz[remember picture] \node[] (srcr) {};\\
\onslide<2->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}
\\\begin{tabular}{p{3.2cm} p{3.2cm}}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
    move  rtmp1 42
    store rstk rtmp1
    cca   rstk -1
    geta  rtmp1 rstk
    cca   rretc 5
    move  rtmp1 pc
    cca   rtmp1 -20
\end{lstlisting}&
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
    load  rtmp1 rtmp1
    cca   rtmp1 -21
    cseal rretd rtmp1 
    move  rretc pc
    xjmp  r1 r2
    cseal rretc rtmp1
    move  rtmp1 0
\end{lstlisting}
\end{tabular}\tikz[remember picture] \node[] (trgr) {};
    \begin{tikzpicture}[remember picture, overlay]
      %White overlay
      \draw<1> node[rectangle,fit=(srccodetl) (srccodebr), fill=white] (whiteoverlay) {};

      % Call red overlay
      \draw<4-> node[rectangle, fit=(calltl) (callbr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (call) {};
      \draw<4-> node[rectangle, fit=(calltl) (callbr)] (calltext) [yshift=-0.1cm] {\textbf{\texttt{call}}};

      % LCM text
      \draw node[right = 3.3cm of trgr, text width = 3cm] {Linear Capability Machine (LCM)};

      % Call stack
      % TODO draw call stack
      \begin{scope}
        \clip<3-> ($(srcr) + (0.1,-1.2)$) rectangle ($(srcr) + (2.1,-.75)$);
        \draw<3-> ($(srcr) + (0.2,-1.5)$) rectangle ($(srcr) + (2,-.75)$) node[pos=.5,align=center] {\scriptsize{...}};
      \end{scope}
      \draw<3-> ($(srcr) + (0.2,-0.75)$) rectangle ($(srcr) + (2,-0.25)$) node[pos=.5,align=center] {\scriptsize{frame 1}};
      \draw<3-> ($(srcr) + (0.2,-0.25)$) rectangle ($(srcr) + (2,0.25)$) node[pos=.5,align=center] {\scriptsize{frame 2}};
      \draw<3-> ($(srcr) + (0.2,.25)$) rectangle ($(srcr) + (2,.75)$) node[pos=.5,align=center] {\scriptsize{frame 3}};

      \draw<3-> ($(srcr) + (1.1,1.25)$) node[align=center] {\footnotesize{Builtin call stack}};
      \draw<3-> ($(srcr) + (0.2,-.75)$) -- ($(srcr) + (0.2,-1)$);
      \draw<3-> ($(srcr) + (2,-.75)$) -- ($(srcr) + (2,-1)$);
      % Return red overlay
      \draw<4-> node[rectangle, fit=(retl) (retr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (ret) {};
      \draw<4-> node[rectangle, fit=(retl) (retr)] (rettext) [yshift=-0.1cm] {\textbf{\texttt{return}}};

      %oLCM text
      \draw<2-> node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics (oLCM)};
%      \draw[dashed] (mid) -- ($(mid) + (7,0)$);
    \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Fully-abstract overlay semantics}
  \centering
  \begin{tikzpicture}
    \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
    \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
    \draw (olcm) edge[bend left, ->,thick] (lcm);
    \draw node[left = 3cm of olcm] (ll1) {};
    \draw node[right = 3cm of olcm] (rl1) {};
    \draw node[below = 1.25cm of ll1] (ll) {};
    \draw node[below = 1.25cm of rl1] (rl) {};
    \draw[dashed] (ll) -- (rl);

    \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
    \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Evaluating the definition}
  \begin{enumerate}
  \item {\itshape Intuitive}\\
    \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
\item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
                                % when reasoning about correctness and security
                                % of programs using \stktokens{}.
  \onslide<3->{Yes, ...}
\item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
  \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
    be used as a target for other compilation phases.}
\item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
  \onslide<5->{Yes, ...}
%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{The \stktokens{} calling convention}
  \begin{onslide}<2->
    ... is explained in the paper.
  \end{onslide}
\end{frame}

\begin{frame}
  \centering
  Thank you!
\end{frame}

% \begin{lstlisting}[basicstyle=\tiny\ttfamily]
%     move rtmp1 42
%     store rstk rtmp1
%     cca rstk -1
%     geta rtmp1 rstk
%     split rstk rretd rstk rtmp1
%     move rtmp1 pc
%     cca rtmp1 -20
%     load rtmp1 rtmp1
%     cca rtmp1 -21
%     cseal rretd rtmp1
%     move rretc pc
%     cca rretc 5
%     cseal rretc rtmp1
%     move rtmp1 0
%     xjmp r1 r2
%     getb rtmp1 rstk
%     minus 1 rtmp1 432
%     move 2 pc
%     cca rtmp2 5
%     jnz rtmp2 rtmp1
%     cca rtmp2 1
%     jmp rtmp2
%     fail
%     splice rstk rdata
%     cca rstk 1
%     move rtmp2 0
% \end{lstlisting}

\end{document}