### 2019-01-12 presentations/illustrations/presentation-illustrations.tex

\documentclass{beamer}
\input{preamble}
\begin{document}

%\input{traditional}
%\input{attacks}

\begin{frame}
  \frametitle{Stack Capabilities Problems}
  \begin{itemize}
  \item Stack capabilities are not a panacea
  \item Stack pointers must not leave stack frame (through memory or stack)
  \item Similar problems for return pointers (omitted)
  \item Solution: local capabilities?
    \begin{itemize}
    \item Make stack and return capabilities local
    \item Cannot leave registers (except on stack) (\sout{Attack 1})
    \item Stack clearing on boundary crossings (\sout{Attack 2})
    \end{itemize}
  \end{itemize}
\end{frame}

%\input{local-prevention}
\input{stktokens-high-level}
%\input{linear-prevention}

\end{document}